### 2022-01-26 chapters/fairness/justicia/conclusion.tex

\section{Discussion and Future Work}
In this paper, we have presented {\framework}, a unified formal framework based on SSAT to verify different fairness metrics in presence of compound protected groups. In our empirical studies, {\framework} achieves higher accuracy and impressive scalability than the existing distribution-based fairness verifiers FairSquare and VeriFair. Moreover, {\framework} verifies different fairness-enhancing algorithms in multiple datasets and classifiers. Additionally, {\framework} is more robust than the sample-based verifier AIF360. As a future work, we incorporate attribute-correlation in the presented learning-based encoding.