### 2022-01-26 chapters/fairness/justicia/table_synthetic.tex

\begin{table}[t!]
    \centering
%        \vspace*{-.2em}
        \setlength{\tabcolsep}{.1em}
            \begin{tabular}{ccccccc}
                \toprule
                Metric & Exact  & {\justicia} & FairSquare & VeriFair & AIF360\\
                 \midrule
				Disparate impact &  $ 0.26 $  &  $ 0.25 $    &  $ 0.99 $  &  $ 0.99 $  &  $ 0.25 $  \\
				Stat. parity &  $ 0.53 $  &  $ 0.54 $    & \textemdash & \textemdash &  $ 0.54 $  \\
                \bottomrule
    \end{tabular}
 \caption{Results on synthetic benchmark.  `\textemdash'~ refers that the verifier cannot compute the metric. }
\label{fairness_justicia_tab:synthetic}
\end{table}