### 2022-01-26 figures/fairness/fvgm/tables/ssat_vs_subset_sum.tex

\begin{figure}
		\begin{center}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_Independent_Adult_LR}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_DAG_Adult_LR}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_time_scatter_plot_Adult_LR}}\\

			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_Independent_Adult_SVM}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_DAG_Adult_SVM}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_time_scatter_plot_Adult_SVM}}\\

			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_Independent_Compas_LR}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_DAG_Compas_LR}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_time_scatter_plot_Compas_LR}}\\

			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_Independent_Compas_SVM}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_DAG_Compas_SVM}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_time_scatter_plot_Compas_SVM}}\\

			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_Independent_Ricci_LR}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_DAG_Ricci_LR}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_time_scatter_plot_Ricci_LR}}\\

			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_Independent_Ricci_SVM}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_DAG_Ricci_SVM}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/ssat_vs_subsetsum_time_scatter_plot_Ricci_SVM}}\\

			\subfloat[]{\includegraphics[scale=0.25]{figures/cactus_subsetsum_ssat_LR_time_.pdf}}
			\subfloat[]{\includegraphics[scale=0.25]{figures/cactus_subsetsum_ssat_SVM_time_.pdf}}

		\end{center}

		\caption{Scalability between {\framework} and Justicia. \red{Each instance represents the runtime of verifying on a dataset and a set of sensitive attributes.}}
\end{figure}