### 2022-01-26 tables/fairness/fvgm/fairness_attacks.tex

\begin{figure}[!t]
	\begin{center}
		\subfloat{\includegraphics[scale=0.24]{figures/disp_fairness_attack}}
		\subfloat{\includegraphics[scale=0.24]{figures/stat_fairness_attack}}\\
	\end{center}
	\caption{Verifying fairness poisoning attack using {\framework}. The red line denotes safety margin, which being exceeded denotes system-vulnerability by the attack algorithm. As the number of poisoned samples increase, disparate impact (DI) decreases and statistical parity (SP)  increases.}
	\label{fig:attack_extended}
\end{figure}