### 2022-01-26 tables/fairness/fvgm/fairness_influence_functions.tex

\begin{figure}[h]
	\begin{center}

		\subfloat[COMPAS]{\includegraphics[scale=0.25]{figures/dependency_exp_Learn-dependency_compas_lr_sex_0}}
		\subfloat[COMPAS]{\includegraphics[scale=0.25]{figures/dependency_exp_Learn-dependency_compas_lr_sex_1}}
		\subfloat[COMPAS]{\includegraphics[scale=0.25]{figures/dependency_exp_Learn-dependency_compas_lr_sex_DI}}\\

		\subfloat[]{\includegraphics[scale=0.25]{figures/dependency_exp_Learn-dependency_adult_lr_sex_0}}
		\subfloat[]{\includegraphics[scale=0.25]{figures/dependency_exp_Learn-dependency_adult_lr_sex_1}}
		\subfloat[]{\includegraphics[scale=0.25]{figures/dependency_exp_Learn-dependency_adult_lr_sex_DI}}\\

	\end{center}
	\caption{Results on computing feature influence functions (FIF) for COMPAS  and Adult dataset.}
\end{figure}