### 2022-01-26 tables/fairness/fvgm/sanity_check.tex

\begin{figure}
	\begin{center}
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_probability}\label{fig:synthetic_dist}}
		\subfloat[]{\includegraphics[scale=0.45]{figures/synthetic_example_BN}\label{fig:synthetic_bn}}

%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_DI_LR}}
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_DI_SVM}}
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_DI_DT}}
%		
%		
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_SPD_LR}}
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_SPD_SVM}}
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_SPD_DT}}
%		
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_max_PPV_LR}\label{fig:synthetic_ppv_max_LR}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_max_PPV_SVM}}\\
%		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_max_PPV_DT}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_min_PPV_LR}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_min_PPV_SVM}\label{fig:synthetic_ppv_min_SVM}}
%		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_min_PPV_DT}\label{fig:synthetic_ppv_b}}

		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_DI_LR}\label{fig:synthetic_sample_size_LR}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_DI_SVM}}\\
%		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_DI_DT}}	
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_SPD_LR}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_SPD_SVM}\label{fig:synthetic_sample_size_SVM}}
%		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_SPD_DT}\label{fig:synthetic_sample_size_b}}

	\end{center}

	\caption{
		Measuring accuracy of different fairness verifiers for Example~\ref{example:intro}. LR, SVM, DT refers to Logistic Regression classifier, Support Vector Machine, and Decision Tree, respectively.}
	\label{fig:synthetic_results}
\end{figure}