### 2022-01-27 chapters/interpretability/relaxed-cnf/example.tex

\section{Illustrative Example}
\label{sec:example}

In this section, we provide a toy example to illustrate  a relaxed-CNF rule which we learn for  Breast Cancer Wisconsin Diagnostic Data Set (WDBC\footnote{\url{https://archive.ics.uci.edu/ml/datasets/Breast+Cancer+Wisconsin+(Diagnostic)}}). The dataset is used to predict whether the cancer is benign or malignant based on the characteristics of an image of  a tumor cell.  {\Framework} generates the following rule for detecting malignant cancer. 
%\vspace{-10pt}

%\begin{comment}
\[
\begin{split}
&[ (  \text{ smoothness }\ge 0.089  + \text{ standard error of area  }\ge 53.78 \\
& + \text{ largest radius }\ge 18.225   )\ge 2  ] +\\
&[   (98.76\le \text{ perimeter } < 114.8  +  \text{ largest smoothness }\ge 0.136 \\
& + 105.95 \le \text{ largest perimeter }  < 117.45     )\ge 2  ]  \ge 1 \end{split}
\]
%\end{comment}

The generated rule has two clauses, each clause consisting of three characteristics of the image of a tumor cell. In this rule, a clause is satisfied if two out of three characteristics match for a test sample. Finally, the rule predicts ``malignant'' cancer if at least one clause is satisfied. The size of this rule is $ 3+3=6 $.  If each clause in this rule is encoded into CNF,  standard encoding technique~\cite{steinke2014pblib} would generate more than one CNF clauses, resulting in higher rule-size. Therefore,   relaxed-CNF rules possesse higher expressibility than  CNF rules. 
We believe that this type of rule is very intuitive and helpful to human experts in decision making.

\begin{comment}
{\scriptsize
	\begin{table}[]
		\setlength{\tabcolsep}{5pt}

		%		%\vspace{-6 pt}
		\begin{center}
			\begin{tabular}{|l | l| }
				%%				\newcommand{\softand}{\ensuremath{\mathsf{soft-AND}}}
				\hline

				\multicolumn{2}{|c|}{ $  \Rule = \ensuremath{\mathsf{softAND}}(\{\text{Test } 1 , \text{Test 
					} 2 \}, 1 $) }  \\
				\hline
				\multicolumn{2}{c}{ }  \\
				\hline
				\multicolumn{1}{|c|}{ Test $ 1 =  \ensuremath{\mathsf{hardOR}} (\cdot ,2 )$} & check \\
				\hline
				$\text{perimeter }> 700.66  $ & \\
				\hline
				$\text{standard error of perimeter }> 91.31 $ & \\
				\hline
				$\text{largest radius }> 23.58   $& \\
				\hline
				$\text{largest texture }> 115.9   $& \\
				\hline
				$\text{largest  concave portions of contour }> 0.15    $& \\
				\hline

				\multicolumn{2}{c}{\multirow{1}{*}{ }} \\
				%				\multicolumn{2}{c}{}\\
				\hline
				\multicolumn{1}{|c|}{ Test $ 2 =  \ensuremath{\mathsf{hardOR}} (\cdot ,2 )$}&  check \\
				\hline
				$ \text{area }> 0.1 $ & \\
				\hline 
				$ \text{largest radius }> 33.65   $ & \\
				\hline
				$ \text{largest perimeter }> 781.18  $&  \\
				\hline

			\end{tabular}
		\end{center}
		%\vspace{-10 pt}
		\caption{Example of relaxed-CNF rules for WDBC dataset.  \ensuremath{\mathsf{hardOR}} takes  a   clause and  $ \eta_l $ as input   and returns $ 1 $  when at least $ \eta_l $ literals are  $ 1 $. \ensuremath{\mathsf{softAND}} takes   a formula and  $ \eta_c $ as input and returns $ 1 $  when at least $ \eta_c $ clauses are  $ 1 $.  Informally,   a test is diagnosed as positive if at least two properties of breast tissue are matched. The patient has malignant cancer if she/he is diagnosed positive in at least one test, otherwise she/he has benign cancer.}
		\label{tab:illustration}
%		\vspace{-40 pt}
	\end{table}
}

\end{comment}