### 2022-01-27 chapters/interpretability/imli/abstract.tex
imli journal paper added

\section{abstract}
Machine learning has become omnipresent with applications in various safety-critical domains such as medical, law, and transportation. In these domains, high-stake decisions provided by machine learning necessitate researchers to design \emph{interpretable} models, where the prediction is understandable to a human. In interpretable machine learning, \emph{rule-based classifiers} are particularly effective in representing the decision boundary through a set of rules comprising input features. Examples of such classifiers include decision trees, decision lists, and decision sets. The interpretability of rule-based classifiers is in general related to the size of the rules, where smaller rules are considered more interpretable. To learn such a classifier, the brute-force direct approach is to consider an optimization problem 
that tries to learn the smallest classification rule that has close to maximum accuracy. This optimization problem is computationally intractable due to its combinatorial nature and thus, the problem is not scalable in large datasets. To this end, in this paper we study the triangular relationship among the accuracy, interpretability, and scalability of learning rule-based classifiers.

The contribution of this paper is an interpretable learning framework {\imli}, that is based on maximum satisfiability (MaxSAT) for synthesizing classification rules expressible in proposition logic. {\imli} considers a joint objective function to optimize the accuracy and the interpretability of classification rules and learns an optimal rule by solving an appropriately designed MaxSAT query. Despite the progress of MaxSAT solving in the last decade, the straightforward MaxSAT-based solution cannot scale to practical classification datasets containing thousands to millions of samples. Therefore, we incorporate an efficient incremental learning technique inside the MaxSAT formulation by integrating mini-batch learning and iterative rule-learning. The resulting framework learns a classifier by iteratively covering the training data, wherein in each iteration, it solves a sequence of smaller MaxSAT queries corresponding to each mini-batch. In our experiments, {\imli} achieves the best balance among prediction accuracy, interpretability, and scalability. For instance, {\imli} attains a competitive prediction accuracy and interpretability w.r.t.\ existing interpretable classifiers and  demonstrates impressive scalability on large datasets where both interpretable and non-interpretable classifiers fail. As an application, we deploy {\imli} in learning popular interpretable classifiers such as decision lists and decision sets.