### 2022-01-30 chapters/fairness/justicia/preliminaries.tex
preliminaries of justicia updated

\section{Background: Fairness and SSAT}
\label{fairness_justicia_sec:preliminaries}

In this section, we define different fairness metrics for a supervised learning problem. Following that, we discuss Stochastic Boolean Satisfiability (SSAT) problem.

\subsection{Fairness Metrics for Machine Learning}\label{fairness_justicia_sec:fairness}

We consider\footnote{{We represent sets/vectors by bold letters, and the corresponding distributions by calligraphic letters. We express random variables in uppercase, and an assignment of a random variable in lowercase.}} a dataset $ \mathbf{D} $ as a collection of triples $ (\nonsensitive, \sensitive, Y) $ generated from an underlying distribution $\mathcal{D}$. $ \nonsensitive \triangleq \{X_1, \dots, X_{\numnonsensitive}\} $ are non-sensitive features whereas $ \sensitive \triangleq \{A_1, \dots, A_{\numsensitive}\} $ are categorical sensitive features.  $Y \in \{0,1\}$ is the binary label (or class) of $(\nonsensitive,\sensitive)$. Each non-sensitive feature $ X_i$ is sampled from a continuous probability distribution {$ \mathcal{X}_i $}, and each sensitive feature $ A_j \in \{0, \dots, N_j\}  $ is sampled from a discrete probability distribution {$ \mathcal{A}_j $}. We use $ (\mathbf{x}, \mathbf{a}) $ to denote the feature-values of  $ (\nonsensitive, \sensitive) $.  For sensitive features, a valuation vector $ \mathbf{a} = [a_1, .., a_{m_2}] $ is called a \textit{compound sensitive group}. For example, consider $ \sensitive = $ \{race, sex\} where race $ \in $ \{Asian, Color, White\} and sex $ \in $ \{female, male\}. Thus $ \mathbf{a} = $ [Asian, female]  is a compound sensitive group. 
We represent a binary classifier trained on the dataset $\mathbf{D}$ as $\alg: (\nonsensitive, \sensitive) \rightarrow \hat{Y} $. Here, $\hat{Y} \in \{0,1\}$ is the predicted class of $ (\nonsensitive, \sensitive) $.

As we illustrated in Example~\ref{fairness_justicia_example:intro}, a classifier $\alg$ that solely optimizes accuracy, i.e., the average number of times $\hat{Y} = Y$, may discriminate certain compound sensitive groups over others~\cite{chouldechova2020snapshot}. In the following, we describe two well-known fairness definitions verified by {\justicia}: group fairness and causal fairness. \red{In each definition, the fairness metric relies on computing the conditional Positive Predictive Value (PPV) of the classifier, denoted by $ \Pr[\hat{Y} = 1 | \mathsf{C}] $, where different fairness metrics are constituted based on the conditional $ \mathsf{C} $.}

\subsubsection{Group Fairness.} Group fairness is categorized into three families: independence, separation and sufficiency, of which {\justicia} verifies independence and separation metrics. 
The independence metrics state that the predicition of the classifier should be independent of compound sensitive groups. Formally, independence notion specifies an equal PPV across all sensitive groups for a classifier $\alg$, i.e., $\Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}]  =  \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}'] , \forall \mathbf{a}, \mathbf{a}' \in \sensitive$.
Since satisfying independence exactly is hard, relaxations of independence fairness metrics, such as \textit{disparate impact} and \textit{statistical parity}~\cite{dwork2012fairness,feldman2015certifying}, are proposed.

\textit{Disparate impact} (DI)~\cite{feldman2015certifying} measures the ratio of PPVs between the most favored group and least favored group, and prescribe it to be close to $1$. Formally, a classifier satisfies $(1 - \epsilon)$-disparate impact if, for $\epsilon \in [0,1] $,
\[
\min_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}]  \ge (1 - \epsilon) \max_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}].
\]
Another popular relaxation of independence metrics  is \textit{statistical parity} (SP) that measures the maximum difference of PPVs among sensitive groups, and prescribe this to be near zero. Formally, an algorithm satisfies $\epsilon$-statistical parity if, for $\epsilon \in [0,1] $, 
\[
\max_{\mathbf{a}}\Pr[\hat{Y} =1 | \mathbf{A} = \mathbf{a}] - \min_{\mathbf{a}}\Pr [\hat{Y} = 1| \mathbf{A} = \mathbf{a}] \le \epsilon.
\]
For both disparate impact and statistical parity, lower value of $\epsilon$ indicates higher group fairness of the classifier.

In the \textit{separation (or classification parity)} notion of fairness, the predicted label $\hat{Y}$ of a classifier is independent of the sensitive features $\sensitive$ given the class labels $Y$. In case of binary classifiers, a popular separation metric is \textit{equalized odds} (EO)~\cite{hardt2016equality} that computes the maximum difference of false positive rates \red{(FPR)} and true positive rates \red{(TPR)} conditioned on sensitive groups.  Lower value of equalized odds indicates better fairness. A classifier satisfies $\epsilon$-equalized odds when the following conditions hold.
\begin{align*}
	&\max_{\mathbf{a}} \Pr[\hat{Y} =1 |\mathbf{A}= \mathbf{a}, Y= 0  ] - 	\min_{\mathbf{a}} \Pr [\hat{Y} = 1|\mathbf{A}= \mathbf{a}, Y = 0] \le \epsilon\\
	&\max_{\mathbf{a}}\Pr[\hat{Y} =1 |\mathbf{A}= \mathbf{a}, Y= 1  ] - 	\min_{\mathbf{a}}  \Pr [\hat{Y} = 1|\mathbf{A}= \mathbf{a}, Y = 1] \le \epsilon
\end{align*}

\subsubsection{Path-specific Causal Fairness.}
Let $ \mathbf{a}_{\max}  \triangleq \argmax_{ \mathbf{a}} \Pr[\hat{Y} =1 |\mathbf{A}=  \mathbf{a}] $. We consider mediator features $ \mediator \subseteq \nonsensitive $ sampled from the conditional distribution $ {\mathcal{Z}_{|\mathbf{A} = \mathbf{a}_{\max}}} $. This emulates the fact that mediator variables have the same sensitive features $ \mathbf{a}_{\max} $.  For $ \epsilon \in [0,1] $,  path-specific causal fairness is defined as 
\[
\max_{\mathbf{a}} \Pr[\hat{Y} = 1 | \sensitive =  \mathbf{a}, \mediator] -  \min_{\mathbf{a}} \Pr[\hat{Y} = 1 | \sensitive = \mathbf{a}, \mediator ] \le \epsilon.
\]

Therefore, PCF constrains that $ \hat{Y} $ is not directly dependent of $ \sensitive $ while $ \sensitive $ may indirectly affects $ \hat{Y} $ only through $ \mediator $. PCF is a variation of counterfactual fairness and causal fairness without mediator features~\cite{bastani2019probabilistic}.

\begin{example}
	Following~\cite{bastani2019probabilistic}, we consider a classifier that decides the hiring of employees based on three features: gender (sensitive), years of experience (non-sensitive), and college-participation (mediator). It is practical to consider that gender $ \in $ \{male, female\} can affect the college-participation of individuals, and all three features are determining factors for the hiring process. Let `male' be the most favored group by the classifier, for instance. Path-specific causal fairness (PCF) ensures that a female candidate should be given a job offer with similar probability as a male candidate (by constraining $ \epsilon \approx 0 $). She,  however,  went to (participated in) college as if she were a male candidate while other non-mediator features such as  `years of experience' are the same.  Therefore, PCF measures the effect of gender on job offer, but ignores the effect of gender on whether candidates went to college.
\end{example}

\subsection{Stochastic Boolean Satisfiability (SSAT)}\label{fairness_justicia_sec:ssat}
Let $\mathbf{B}  = \{\bool_1, \dots, \bool_m\}  $ be a set of Boolean variables. A \textit{literal} is a variable $ \bool_i $ or its complement $ \neg \bool_i $. 
A propositional formula $\phi$ defined over $\mathbf{B}$ is in \textit{Conjunctive Normal Form (CNF)} if $\phi$   is  a conjunction of clauses and each clause is a disjunction of literals. 
%\red{DNF (disjunctive normal form) is the  complement of CNF where  the formula is a disjunction of clauses and each clause is  a conjunction of literals.} 
Let $ \sigma $ be an assignment to the  variables $ \bool_i \in \mathbf{B} $  such that $ \sigma(\bool_i) \in \{1, 0\} $. The propositional  \textit{satisfiability} problem (SAT) finds an assignment $ \sigma $ to all $ \bool_i \in  \mathbf{B} $ such that the formula $ \phi $ is evaluated to be $1$ (equivalently, true). 
In contrast to the SAT problem, the \textit{Stochastic Boolean Satisfiability} (SSAT) problem~\cite{littman2001stochastic} is computes the  probability of the satisfaction of the formula $\phi$ defined on \textit{quantified} Boolean variables. 
An SSAT formula is of the form
\begin{equation}\label{fairness_justicia_eq:ssat}
\Phi = Q_1\bool_1, \dots, Q_m \bool_m,\; \phi, 
\end{equation}
where $ Q_i \in \{\exists, \forall, \R^{p_i}\} $ is either of the existential ($\exists$), universal ($\forall$), or randomized ($\R^{p_i}$) quantifiers on $\bool_i$ and $\phi$ is a quantifier-free CNF formula. In case of a randomized quantifier $ \R^{p_i} $, $ p_i \in [0,1] $ is the probability of $ \bool_i $ being assigned to $ 1 $. In the SSAT formula $ \Phi $, the quantifier part $ Q_1\bool_1, \dots, Q_m \bool_m $ is known as the \textit{prefix} of the formula $ \phi $.  Let $ \bool $ be the outermost variable in the prefix. The semantics of SSAT formulas are defined recursively in the following.
\begin{enumerate}
	\item $ \Pr[\text{true}] = 1 $,  $ \Pr[\text{false}] = 0 $, 
	\item $ \Pr [\Phi] = \max_{\bool} \{\Pr[\Phi|_{\bool}], \Pr[\Phi|_{\neg \bool}]\}$ if $ \bool $ is existentially quantified ($ \exists $), 
	\item $ \Pr [\Phi] = \min_{\bool} \{\Pr[\Phi|_{\bool}], \Pr[\Phi|_{\neg \bool}]\} $ if $ \bool $ is universally quantified ($ \forall $), 
	\item $ \Pr [\Phi] = p\Pr[\Phi|_{\bool}] + (1-p) \Pr[\Phi|_{\neg \bool}] $ if $ \bool $ is randomized quantified ($\R^{p}$) with probability $p$ of being $\text{true}$,
\end{enumerate}
where $ \Phi|_{\bool} $ and $ \Phi|_{\neg \bool} $ denote the SSAT formulas derived by eliminating the outermost quantifier of $ \bool $  by substituting the value of $ \bool $ in the formula $ \phi $ with $ 1 $ and $ 0 $, respectively. In this paper, we focus on two specific types of SSAT formulas:  \textit{random-exist} (RE) SSAT and \textit{exist-random} (ER) SSAT. In the ER-SSAT (resp.\ RE-SSAT) formula, all existentially (resp.\ randomized) quantified variables are followed by randomized (resp.\ existentially) quantified variables in the prefix.

\begin{remark}
	ER-SSAT problem is $\mathrm{NP}^{\mathrm{PP}}$-hard whereas RE-SSAT problem is $\mathrm{PP}^{\mathrm{NP}}$-complete~\cite{littman2001stochastic}.
\end{remark}

The problem of SSAT and its variants have been pursued by theoreticians and practitioners for over three decades~\cite{majercik2005dc,fremont2017maximum,huang2006combining}. We refer the reader to~\cite{lee2017solving,lee2018solving} for detailed survey. It is worth remarking that the past decade has witnessed a significant performance improvements of SSAT solving, thanks to the close integration of techniques from SAT solving with advances in weighted model counting~\cite{sang2004combining,chakraborty2013scalable,chakraborty2014distribution}.

\noindent\makebox[\linewidth]{\rule{\paperwidth}{2.5pt}}

%In these extreme situations, the reduced problem is NP-hard and PP-hard respectively.
%We can show solving ER- and RE-SSAT are equivalent to solving an E-MAJSAT/e-threshold SAT problem. Thus it is $NP^{PP}$ hard. 
%Equivalent formulation of this is computing MAP in a belief network.

\begin{comment}
	The Stochastic Boolean Satisfiability problem (SSAT) allows existential ($ \exists $) and randomized ($ \R $) quantifiers over the Boolean variables and evaluates the probability of satisfaction of the CNF formula.  Based on the order of quantified variables, an SSAT instance can be either a random-exists (\textbf{RE}) SSAT  formula or an exist-random (\textbf{ER}) SSAT formula.

	Let $ \phi(E,R) $ be a CNF formula over Boolean variables $ E \cup R $ where each $ e_i \in E $ is an exist quantified variable and each $ r_j \in R $ is a randomized quantified variable.  For each $ r_j $, $ p_j = \Pr(\sigma(r_j) = 1) $ denotes the probability of $ r_j $ assigning to true. Now an \textbf{RE}-SSAT formula is defined as

	\[
	\R^{p_1} r_1, \dots, \R^{p_m} r_m, \exists e_1, \dots, \exists e_n \; \phi (E,R)
	\]

	where $ |R| = m,  |E| = n  $. Intuitively,  for the random assignment of $ r_1, \dots, r_m $ there exists assignments of $ e_1, \dots, e_n $, for which, we ask the \red{maximum}  probability of satisfaction of $ \phi $. Similarly, an \textbf{ER}-SSAT formula is defined as 
	\[
	\exists e_1, \dots, \exists e_n, \R^{p_1} r_1, \dots, \R^{p_m} r_m, \; \phi (E,R). 
	\]

	is to evaluate a Boolean formula containing both existential and randomized quantifiers~\cite{littman2001stochastic}.  For example:

	\[
	\exists x_1, \R^{p_2} y_2, \exists x_3, \dots, \R^{p_n} y_n (E[\phi (\mathbf{x})]\ge \theta)
	\]

	In words, this formula asks whether there is a value for $ x_1 $ such that for random values of $ y_2 $ (choose $ y_2 = 1 $ with probability $ p_2 $), there exists a value of $ x_2 $, such that$ \dots $ the expected value of the Boolean formula $ \phi(\mathbf{x}) $ meets or exceeds a threshold $ 0 \le \theta \le 1 $. Here variables $ x_i, y_i \in \mathbf{x} $ are Boolean.

	While the above definition of SSAT has alternating quantifiers, it is not a strict restriction. One could first place random quantifiers followed by exists quantifiers or vice versa. Moreover, the threshold $ \theta $ can be omitted from the definition where the question is to find out the probability of $ \phi(\mathbf{x}) =1 $ given the quantifiers over Boolean variables $ x_i $ and $ y_i $.

\end{comment}