### 2022-04-07 chapters/fairness/fvgm/abstract.tex
updated files

\begin{abstract}
In recent years, machine learning (ML) algorithms have been deployed in safety-critical and high-stake decision-making, where the \textit{fairness} of algorithms is of paramount importance. Fairness in ML  centers on detecting bias towards certain demographic populations induced by an ML classifier and proposes algorithmic solutions to mitigate the bias with respect to different fairness definitions.  To this end, several \textit{fairness verifiers} have been proposed that compute the bias in the prediction of an ML classifier\textemdash essentially beyond a finite dataset\textemdash given the probability distribution of input features. In the context of verifying \textit{linear classifiers}, existing fairness verifiers are limited by \emph{accuracy} due to imprecise modeling of correlations among features and \emph{scalability} due to restrictive formulations of the classifiers as SSAT/SMT formulas or by sampling.

In this paper, we propose an efficient fairness verifier, called \fvgm, that encodes the correlations among features as a Bayesian network. In contrast to existing verifiers, \fvgm~proposes a \textit{stochastic subset-sum} based approach for verifying linear classifiers. Experimentally, we show that  \fvgm~leads to an \textit{accurate} and \textit{scalable} assessment for more diverse families of fairness-enhancing algorithms, fairness attacks, and group/causal fairness metrics than the state-of-the-art fairness verifiers. We also demonstrate that {\fvgm} facilitates the computation of fairness influence functions as a stepping stone to detect the source of bias induced by subsets of features.
\end{abstract}

%that encodes the correlations among features as a Bayesian network and \red{enables disentangled measurement of bias induced by the data and the classifier}.