### 2022-04-07 chapters/fairness/fvgm/background.tex

	\section{Background}
	In this section, we define different fairness metrics proposed for classification. Following that, we state basics of stochastic subset sum and Bayesian networks that are the main components of our methodology.

	\textbf{Fairness in ML.}
	We consider\footnote{{We represent sets/vectors by bold letters, and the corresponding distributions by calligraphic letters. We express random variables in uppercase, and an assignment of a random variable in lowercase.}} a dataset $ \mathbf{D} $ as a collection of triples $ (\nonsensitive, \sensitive, Y) $ generated from an underlying distribution $\mathcal{D}$. $ \nonsensitive \triangleq \{X_1, \dots, X_{m_1}\} $ are non-sensitive features whereas $ \sensitive \triangleq \{A_1, \dots, A_{m_2}\} $ are categorical sensitive features.  $Y \in \{0,1\}$ is the binary label (or class) of $(\nonsensitive,\sensitive)$. Each non-sensitive feature $ X_i$ is sampled from a continuous probability distribution {$ \mathcal{X}_i $}, and each sensitive feature $ A_j \in \{0, \dots, N_j\}  $ is sampled from a discrete probability distribution {$ \mathcal{A}_j $}. 
%	\red{Therefore, the distribution $\mathcal{D}$ is the joint distribution on $(\nonsensitive,\sensitive)$, where  $\mathcal{X}_i$ and $\mathcal{A}_i$ are the marginal distributions.} 
	%Therefore, $ \mathcal{D} = \prod_{i=1}^{m} \mathcal{X}_i \prod_{j=1}^{n} \mathcal{A}_j $ denotes the product distribution of sensitive and non-sensitive features and thus considered the underlying distribution for the dataset $ D $.  Additionally, $ Y \in \{0,1\} $ denotes the binary class label of $ (\nonsensitive, \sensitive) $. Let $ \alg : (\nonsensitive, \sensitive) \rightarrow \hat{Y} $  be a binary classier trained on $ D $, where $ \hat{Y} \in \{0,1\} $ is the predicted class. The positive predictive value (PPV) of $ \alg $, denoted as $ \Pr[\hat{Y} = 1] $, is the probability of positive outcome.
 	We use $ (\mathbf{x}, \mathbf{a}) $ to denote the feature-values of  $ (\nonsensitive, \sensitive) $.  For sensitive features, a valuation vector $ \mathbf{a} = [a_1, .., a_{m_2}] $ is called a \textit{compound sensitive group}. For example, consider $ \sensitive = $ \{race, sex\} where race $ \in $ \{Asian, Color, White\} and sex $ \in $ \{female, male\}. Thus $ \mathbf{a} = $ [Asian, female]  is a compound sensitive group. 
 	We represent a binary classifier trained on the dataset $\mathbf{D}$ as $\alg: (\nonsensitive, \sensitive) \rightarrow \hat{Y} $. Here, $\hat{Y} \in \{0,1\}$ is the predicted class of $ (\nonsensitive, \sensitive) $.

 	We now discuss different fairness metrics in the literature based on the prediction of a classifier~\cite{feldman2015certifying,hardt2016equality,nabi2018fair}.  	
 	In this paper, {\fvgm} verifies two families of fairness metrics: group fairness (first three in the following) and path-specific causal fairness.

 	\begin{enumerate}[leftmargin=*]
 		\itemsep0em 
 		\item \textit{Disparate Impact} (DI): A classifier  satisfies $(1 - \epsilon)$-disparate impact if for $\epsilon \in [0,1] $,
 		$
 		\min_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}]  \ge (1 - \epsilon) \max_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}].
 		$
 		\item \textit{Statistical Parity} (SP): A classifier satisfies $\epsilon$-statistical parity if for $\epsilon \in [0,1] $, 
 		$
 		\max_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}] - \min_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}] \le \epsilon.
 		$
 		\item \textit{Equalized Odds} (EO): 	A classifier satisfies $\epsilon$-equalized odds if for $\epsilon \in [0,1] $,
 		$ \max_{\mathbf{a}}\Pr[\hat{Y} =1 |\mathbf{A}= \mathbf{a}, Y= 0  ] - \min_{\mathbf{a}}\Pr [\hat{Y} = 1|\mathbf{A}= \mathbf{a}, Y = 0] \le \epsilon, $ and $
 		\max_{\mathbf{a}}\Pr[\hat{Y} =1 |\mathbf{A}= \mathbf{a}, Y= 1  ] - \min_{\mathbf{a}}\Pr [\hat{Y} = 1|\mathbf{A}= \mathbf{a}, Y = 1] \le \epsilon.
 		$
 		\item \textit{Path-specific Causal Fairness} (PCF): 
 		Let $ \mathbf{a}_{\max}  \triangleq \argmax_{ \mathbf{a}} \Pr[\hat{Y} =1 |\mathbf{A}=  \mathbf{a}] $. We consider mediator features $ \mediator \subseteq \nonsensitive $ sampled from the conditional distribution $ {\mathcal{Z}_{|\mathbf{A} = \mathbf{a}_{\max}}} $. This emulates the fact that mediator variables have the same sensitive features $ \mathbf{a}_{\max} $.  For $ \epsilon \in [0,1] $,  path-specific causal fairness is defined as 
 		$
 		\max_{\mathbf{a}}\Pr[\hat{Y} = 1 | \sensitive =  \mathbf{a}, \mediator] - \min_{\mathbf{a}}\Pr[\hat{Y} = 1 | \sensitive = \mathbf{a}, \mediator ] \le \epsilon
 		$.
 	\end{enumerate}

 	  For all of the above metrics, lower value of $\epsilon$ indicates higher fairness demonstrated by the classifier $\mathcal{M}$. Following the observation of~\cite{ghosh2020justicia},  computing all of the aforementioned fairness metrics is equivalent to computing the maximum and minimum of the \textit{Positive Predictive Value} (PPV) of the classifier, denoted as $\Pr[\hat{Y}=1|\mathbf{A} =\mathbf{a}]$, for all compound sensitive groups $\mathbf{a}$ from $ \sensitive $. Thus, in Section~\ref{fvgm_sec:fvgm}, we focus on computing the maximum and minimum PPVs and then extend it to assess corresponding fairness metrics. We call the group for which PPV is maximum (minimum) as the \textit{most (least) favored group}.

% 	\textit{Disparate impact} (DI)~\cite{feldman2015certifying} measures the ratio of PPVs between the most favored group and least favored group, and prescribe it to be close to $1$. Formally, 
% 	Another popular relaxation of independence metrics  is \textit{statistical parity} (SP) that measures the difference of PPV among sensitive groups, and prescribe this to be near zero. Formally, an algorithm satisfies $\epsilon$-statistical parity if, for $\epsilon \in [0,1] $, 
% 	\[
% 	\max_{\mathbf{a}_i, \mathbf{a}_j \in A}|\Pr[\hat{Y} =1 | A = \mathbf{a}_i] - \Pr [\hat{Y} = 1| A = \mathbf{a}_j]| \le \epsilon.
% 	\] 	
% 	
% 	In the \textit{separation (or classification parity)} notion of fairness, the predicted label $\hat{Y}$ of a classifier $\alg$ is independent of the sensitive features $A$ given the actual class labels $Y$. In case of binary classifiers, a popular separation metric is \textit{equalized odds} (EO)~\cite{hardt2016equality} that computes the difference of false positive rates (FPR) and the difference of true positive rates (TPR) among all compound sensitive groups. 
% 	Lower value of equalized odds indicates better fairness.
% 	A classifier $\alg$ satisfies $\epsilon$-equalized odds if, for all compound sensitive groups $\mathbf{a}, \mathbf{a} \in A$,
% 	\begin{equation}
% 	\begin{split}
% 	|\Pr[\hat{Y} =1 | A = \mathbf{a}, Y= 0  ] - \Pr [\hat{Y} = 1| A = \mathbf{a}, Y = 0]| &\le \epsilon, \\ 
% 	|\Pr[\hat{Y} =1 | A = \mathbf{a}, Y= 1  ] - \Pr [\hat{Y} = 1| A = \mathbf{a}, Y = 1]| &\le \epsilon.\notag
% 	\end{split}
% 	\end{equation}
% 	
% 	Path-specific causal fairness (PCF) states that the outcome $ \hat{Y} $ should not depend directly on a sensitive feature, but may depend indirectly on the sensitive feature through other \textit{mediator} features[\cite{?}]. 
% 	
% 	
% 	
% 	\red{Need to clarify, group fairness and causal fairness?}
% 	
% 	
% 	\blue{It may be space consuming to discuss different fairness metrics, in which case we can mention that computing PPV is the core component of estimating different fairness metrics.}
%

	\textbf{Stochastic Subset Sum Problem ({\stochastic}).} 
	Let $ \mathbf{B} \triangleq \{B_i\}_{i=1}^{|\mathbf{B}|}$ be a set of Boolean variables and $ w_i \in \mathbb{Z} $ be the weight of $ B_i $. Given a constraint of the form  $\sum_{i = 1}^ {|\mathbf{B}|} w_i B_i = \tau $, for a constant threshold $ \tau \in \mathbb{Z} $, the subset-sum problem seeks to compute an assignment $\mathbf{b} \in \{0,1\}^{|\mathbf{B}|}$ such that the constraint evaluates to true when $\mathbf{B}$ is substituted with $\mathbf{b}$. Subset sum problem is known to be a $ \mathrm{NP} $-complete problem and well-studied in theoretical computer science~\cite{kleinberg2006algorithm}. The \textit{counting} version of the subset-sum problem counts all $ \mathbf{b} $'s for which the above constraint holds; and this problem belongs to the complexity class $ \mathrm{\#P} $. In this paper, we consider the counting problem for the constraint $\sum_{i = 1}^ {|\mathbf{B}|} w_i B_i \ge \tau $ where variables $ B_i $'s are stochastic. More precisely, we define a \textit{stochastic subset-sum} problem, namely {\stochastic}, that computes $ \Pr[\sum_{i = 1}^ {|\mathbf{B}|} w_iB_i \ge \tau] $.    Details of {\stochastic} are in Section~\ref{fvgm_sec:stochastic_sum_set_sum}.

	\textbf{Bayesian Network.}
	In general, a Probabilistic Graphical Model~\cite{koller2009probabilistic}, and specifically a \textit{Bayesian network}~\cite{pearl1985bayesian,chavira2008probabilistic}, encodes the dependencies and conditional independence between a set of random variables. In this paper, we leverage an access to a Bayesian network on $ \nonsensitive \cup \sensitive $ that represents the joint distribution on them. 	A Bayesian network is denoted by a pair $ (\graph, \theta)$, where $ \graph \triangleq (\mathbf{V}, \mathbf{E}) $ is a DAG (Directed Acyclic Graph), and $\theta$ is a set of parameters encoding the conditional probabilities induced by the joint distribution under investigation. Each vertex $V_i \in \mathbf{V}$ corresponds to a random variable. Edges $ \mathbf{E} \in \mathbf{V} \times \mathbf{V} $ imply conditional dependencies among variables. For each variable $ V_i \in \mathbf{V} $, let $ \parent(V_i) \subseteq \mathbf{V} \setminus \{V_i\} $ denote the set of parents of $ V_i $. Given $\parent(V_i)$ and parameters $\theta$, $ V_i $ is independent of its other non-descendant variables in $\graph$. Thus, for the assignment $ v_i $ of $ V_i $ and $ \mathbf{u} $ of $ \parent(V_i) $, the aforementioned semantics of a Bayesian network encodes the joint distribution of $\mathbf{V}$ as:

	\begin{equation}
		\begin{split}
				\Pr[V_1=v_1, \dots, &V_{|\mathbf{V}|}=v_{|\mathbf{V}|}] = \\ 
			&\prod_{i=1}^{|\mathbf{V}|} \Pr[V_i = v_i | \parent(V_i) = \mathbf{u}; \theta].
		\end{split}
	\label{fvgm_eq:BN}
	\end{equation}

	\iffalse	
	$ \factors $ contains a factor $ \factor $ representing a function over the  assignment or valuation of  $ \{V_i\} \cup \parent(V_i)  $. Let $ v $ and $ \mathbf{u} $ denote an assignment of $ V_i $ and $ \parent(V_i) $, respectively. 	Formally a factor $ \factor(V_i = v, \parent(V_i) = \mathbf{u}) \triangleq  \Pr[V_i = v | \parent(V_i) = \mathbf{u}] $ denotes the conditional probability of $ V_i = v $ given $ \parent(V_i) = \mathbf{u} $. Avoiding notation clutter, we drop $ v $ and $ \mathbf{u} $ and use $ \factor(V_i, \parent(V_i)) \triangleq  \Pr[V_i| \parent(V_i)] $, when it is clear from the context.
	\fi