### 2022-04-07 chapters/fairness/fvgm/framework2.tex

\input{chapters/fairness/fvgm/dp_illustration}	
\subsection{{\stochastic} with Correlated Variables} 
\label{fvgm_sec:dp_with_BN}
In {\stochastic} presented in Section~\ref{fvgm_sec:stochastic_sum_set_sum}, we consider all  Boolean variables to be probabilistically independent. This independence assumption often leads to an \textit{inaccurate estimate} of PPVs because both sensitive and non-sensitive features can be correlated in practical fairness problems. Therefore, we extend {\stochastic} to include correlations among variables.

We consider a Bayesian network $ \BN = (\graph, \theta) $ to represent correlated variables, where $ G \triangleq (\mathbf{V}, \mathbf{E}) $, $ \mathbf{V} \subseteq \mathbf{B} $, $ \mathbf{E} \subseteq \mathbf{V} \times \mathbf{V}  $, and $ \theta $ is the parameter of the network.  In  $ \BN $, we constrain that there is no conditional probability of choice (i.e., existential and universal) variables as we optimize their assignment in {\stochastic}. Choice variables, however, can impose conditions on chance (i.e., random) variables. In practice, we achieve this by allowing no incoming edge on choice variables while learning $ \BN $ (ref. Section~\ref{fvgm_sec:experiments}).

For a chance variable $ B_i \in \mathbf{V} $, let $ \parent(B_i) $ denote its parents. According to Eq.~\eqref{fvgm_eq:BN},  for an assignment $ \mathbf{u} $ of $ \parent(B_i) $, $ \BN $ ensures $ B_i $ to be independent of other non-descendant variables in $ \mathbf{V} $. Hence, in the recursion of Eq.~\eqref{fvgm_eq:dp_recurse}, we substitute  $ p_i $  with  $ \Pr[B_i = 1| \parent(B_i) = \mathbf{u}] $. In order to explicate the dependence on $ \mathbf{u} $, we denote the expected solution of $ S(\mathbf{B}[i:n], \tau) $ as 
$ \mathsf{dp}(i, \tau, \mathbf{u}) $, which for $ B_i \in \mathbf{V} $ is modified as follows:

\begin{align*}
	\mathsf{dp}(i,  &\tau, \mathbf{u}) = \Pr[B_i = 0| \parent(B_i) = \mathbf{u}] \mathsf{dp}(i+1, \tau, \mathbf{u} \cup \{0\})  \\
	& + \Pr[B_i = 1| \parent(B_i) = \mathbf{u}]  \mathsf{dp}(i+1, \tau-w_i, \mathbf{u} \cup \{1\}).
\end{align*}

Since $ \mathsf{dp}(i, \tau, \mathbf{u}) $ involves  $ \mathbf{u} $, we initially perform a topological sort of $ \mathbf{V} $ to know the assignment of parents before computing $ \mathsf{dp} $ on the child. Moreover, there are $ 2^{|\parent(B_i)|} $ assignments of $ \parent(B_i) $, and we compute $ \mathsf{dp}(i, \tau, \mathbf{u}) $ for $ \mathbf{u} \in \{0,1\}^{|\parent(B_i)|} $ to incorporate all conditional probabilities into $ \stochastic $.  For this enumeration, we do not store $ \mathsf{dp}(i, \tau, \mathbf{u}) $ in memory. However, for $ B_i \not \in \mathbf{V} $ that does not appear in the network, we instead compute $ \mathsf{dp}(i, \tau) $ and store it in memory as in Section~\ref{fvgm_sec:dp_formulation}, because $ B_i $ is not correlated with other variables.  Lemma~\ref{fvgm_lm:complexity_dp_with_bn} presents the complexity of solving {\stochastic} with correlated variables, wherein unlike Lemma~\ref{fvgm_lemma:complexity_sss}, the  complexity differentiates based on variables in $ \mathbf{V} $ (exponential) and $ \mathbf{B}\setminus \mathbf{V} $ (pseudo-polynomial).

\begin{lemma}
	\label{fvgm_lm:complexity_dp_with_bn}
	Let $ \mathbf{V} \subseteq \mathbf{B} $ be the set of vertices in the Bayesian network and $ n'' $ be the number of existential and universal variables in $ \mathbf{B} \setminus \mathbf{V} $. Let $ w'_{\exists} = \sum_{B_i \in \mathbf{B} \setminus \mathbf{V} | q_i = \exists} \max\{w_i, 0\}$  and $ w'_{\forall} = \sum_{B_i \in \mathbf{B} \setminus \mathbf{V} | q_i = \forall} \min\{w_i, 0\}$ be the sum of considered weights of existential and universal variables, respectively that only appear in $ \mathbf{B} \setminus \mathbf{V} $. To exactly compute {\stochastic} with correlated variables in the dynamic programming approach,  time complexity is $ \mathcal{O}(2^{|\mathbf{V}|} + (n - n'' - |\mathbf{V}|)(\tau + |w_{neg}| - w'_{\exists} - w'_{\forall}) + n'') $ and space complexity is $ \mathcal{O}((n - n'' - |\mathbf{V}|)(\tau + |w_{neg}| - w'_{\exists} - w'_{\forall})) $.
\end{lemma}

\textbf{A Heuristic for Faster Computation.} We observe that to encode conditional probabilities, we enumerate all assignments of variables in $ \mathbf{V} $. For computing PPVs of a linear classifier with correlated features, we consider a heuristic to sort variables in $ \mathbf{B} = \sensitive \cup \nonsensitive $. Let $ \mathbf{V} \subseteq \mathbf{B} $ be the set of vertices in the network and $ \mathbf{V}^c = \mathbf{B} \setminus \mathbf{V} $. In this heuristic, we sort sensitive variables $ \sensitive $ by positioning $ \sensitive \cap \mathbf{V} $ in the beginning followed by $ \sensitive \cap \mathbf{V}^c $. Then we order the variables $ \mathbf{B} $ such that variables in $ \sensitive $ precedes those in $ \nonsensitive \cap \mathbf{V} $, and the variables in $ \nonsensitive \cap \mathbf{V}^c $ follows the ones in $ \nonsensitive \cap \mathbf{V} $. This sorting allows us to avoid repetitive enumeration of variables in $ \mathbf{V} \subseteq \mathbf{B} $ as they are placed earlier in $ \mathbf{B} $.

	\begin{example}
		We extend Example~\ref{fvgm_example:subset-sum} with a Bayesian Network $ (\graph, \theta) $ with $ \mathbf{V} = \{P, Q\} $ and $ \mathbf{E} = \{(P,Q)\} $. Parameters $ \theta $ imply conditional probabilities $ \Pr[Q|P] = 0.6 $ and $ \Pr[Q|\neg P] = 0.3 $. 	In Figure~\ref{fvgm_fig:example_BN}, we enumerate all  assignment of $ P $ and $ Q $ to  incorporate all conditional probabilities of $ Q $ given $ P $. We, however, observe that the dynamic programming solution in Section~\ref{fvgm_sec:dp_formulation} still prunes search space for variables that do not appear in $ \mathbf{V} $, such as $ \{R, S\} $. Hence following the calculation in Figure~\ref{fvgm_fig:example_BN}, we obtain the maximum PPV  as $ 0.65 $ for $ P = 1 $. The minimum PPV (not shown)  is similarly calculated as $ 0.11 $ for $ P = 0 $. 
	\end{example}

	\subsection{Fairness Verification with Computed PPVs} 
	Given a classifier $\mathcal{M}$, a  distribution $\mathcal{D}$, and a fairness metric $f$, verifying whether a classifier is $\epsilon$-fair ($\epsilon \in [0,1]$) is equivalent to computing $\mathds{1}[f(\mathcal{M}|\mathcal{D})\leq \epsilon]$. We now compute $f(\mathcal{M}|\mathcal{D})$ based on the maximum PPV $ \max_{ \mathbf{a}} \Pr[\hat{Y} =1 | \sensitive = \mathbf{a}] $ and the  minimum PPV $ \min_{ \mathbf{a}} \Pr[\hat{Y} =1 | \sensitive = \mathbf{a}] $ of a classifier.

	For measuring fairness metric SP, we compute the difference $ \max_{ \mathbf{a}} \Pr[\hat{Y} =1 | \sensitive = \mathbf{a}]  - \min_{ \mathbf{a}} \Pr[\hat{Y} =1 | \sensitive = \mathbf{a}] $. We, however, deploy {\fvgm} twice while measuring EO: one for the distribution $ \mathcal{D} $ conditioned on $ Y = 1  $ and another for $ Y = 0 $. 
	In each case, we compute $ \max_{ \mathbf{a}} \Pr[\hat{Y} =1 | \sensitive = \mathbf{a}, Y = y ]  - \min_{ \mathbf{a}} \Pr[\hat{Y} =1 | \sensitive = \mathbf{a}, Y = y ] $ for $ y \in \{0,1\} $ and take the  maximum difference as the value of EO.  
	For measuring causal metric PCF, we compute  $ \max_{ \mathbf{a}} \Pr[\hat{Y} =1 | \sensitive = \mathbf{a}, \mathbf{Z}] $ and  $ \min_{ \mathbf{a}} \Pr[\hat{Y} =1, \mathbf{Z}| \sensitive = \mathbf{a} , \mathbf{Z}] $ conditioned on mediator features $ \mathbf{Z} $ and take their difference. 
	To measure DI, we compute the ratio $ \max_{ \mathbf{a}} \Pr[\hat{Y} =1 | \sensitive = \mathbf{a}] / \min_{ \mathbf{a}} \Pr[\hat{Y} =1 | \sensitive = \mathbf{a}] $. In contrast to other fairness metrics, DI closer to 1 indicates higher fairness level. Thus, we verify whether a classifier achieves $(1 - \epsilon)$-DI by checking $ \mathds{1}[DI(\alg| \mathcal{D}) \ge 1 - \epsilon] $.

	\subsection{Extension to Practical Settings}\label{fvgm_sec:practical}
	For verifying linear classifiers with real-valued features and coefficients, we preprocess them so that {\fvgm} can be invoked. Let $ X_c \in \mathbb{R} $ be a continuous real-valued feature with coefficient $ w_c \in \mathbb{R} $ in the classifier. We discretize $ X_c $ to a set $ \mathbf{B}_{c} $ of $ k $ Boolean variables using binning-based discretization and assign a Boolean variable to each bin. Hence, $ B_i \in \mathbf{B}_{c} $ becomes $ 1 $, when $ X_c $ belongs to the $ i^\text{th} $ bin. Let $ \mu_i $ denote the mean of feature-values within $ i^\text{th} $ bin. We then set the coefficient of $ B_i $ as $ w_c\mu_i $. By law of large numbers, $ X_c \approx \sum_i \mu_iB_i $ for infinitely many bins~\cite{grimmett2020probability}. Finally, we multiply the coefficients of discretized variables by $ l \in \mathbb{N} \setminus \{0\} $ and round to an integer. Accuracy of the preprocessing step relies on the number of bins $ k $ and the multiplier $ l $. Therefore, we empirically fine-tune both $ k $ and $ l $ by comparing  the processed classifier with the initial classifier on a validation dataset.
\begin{comment}
	\begin{figure}
		\centering
		\subfloat[Exploration tree of stochastic sub-set sum problem.]{	\includegraphics[scale=0.3]{figures/exploration_tree_DP}\label{fvgm_fig:example_dp}}
		\subfloat[Bayesian network]{\includegraphics[scale=.4]{figures/example_BN}\label{fvgm_fig:example_BN}}
		\subfloat[Exploration tree of stochastic sub-set sum problem with Bayesian network]{\includegraphics[scale=.4]{figures/exploration_tree_DP_BN}	\label{fvgm_fig:example_dp_BN}
		}

	\end{figure}
\end{comment}