### 2022-05-27 chapters/list_of_papers.tex

%\clearpage
%\begin{center}
%	\large\textbf{List of Publications}
%	\normalsize
%\end{center}

\section*{List of Publications}
This thesis is based on the following publications.

\begin{enumerate}
	\item \href{https://arxiv.org/pdf/2205.06936.pdf}{\blue{Efficient Learning of Interpretable Classification Rules}} \\
	\underline{Bishwamittra Ghosh}, Dmitry Malioutov, Kuldeep S. Meel\\
	In Proceedings of JAIR, 2022

	\item \href{https://arxiv.org/pdf/2109.09447.pdf}{\blue{Algorithmic Fairness Verification with Graphical Models}} \\
	\underline{Bishwamittra Ghosh}, Debabrota Basu, Kuldeep S. Meel\\
	In Proceedings of AAAI, 2022

	\item \href{https://arxiv.org/pdf/2009.06516.pdf}{\blue{Justicia: A Stochastic SAT Approach to Formally Verify Fairness}} \\
	\underline{Bishwamittra Ghosh}, Debabrota Basu, Kuldeep S. Meel\\
	In Proceedings of AAAI, 2021

	\item \href{https://bishwamittra.github.io/publication/ecai_2020/paper.pdf}{\blue{Classification Rules in Relaxed Logical Form
	}} \\
	\underline{Bishwamittra Ghosh}, Dmitry Malioutov, Kuldeep S. Meel\\
	In Proceedings of ECAI, 2020

	\item \href{https://bishwamittra.github.io/publication/imli-ghosh.pdf}{\blue{IMLI: An Incremental Framework for MaxSAT-Based Learning of Interpretable Classification Rules}}\\
	\underline{Bishwamittra Ghosh}, Kuldeep S. Meel\\
	In Proceedings of AIES, 2019

\end{enumerate}
\clearpage