### 2022-06-01 chapters/fairness/fif/abstract.tex
adding fifs and introduction

\begin{abstract}

	Fairness in machine learning has attained significant focus due to the widespread application of machine learning in high-stake decision-making tasks. Unless regulated with a fairness objective, machine learning classifiers might demonstrate unfairness/bias towards certain demographic populations in the data. Thus, the quantification and mitigation of the bias induced by classifiers have become a central concern. In this paper, \textit{we aim to quantify the influence of different features on the bias of a classifier}. To this end, we propose a framework of \emph{Fairness Influence Function} (FIF), and compute it as a scaled difference of conditional variances in the classifier's prediction. We also instantiate an algorithm, {\fairXplainer}, that uses variance decomposition among the subset of features and a local regressor to compute FIFs accurately, while also capturing the intersectional effects of the features. Our experimental analysis validates that {\fairXplainer} captures the influences of both individual features and higher-order feature interactions, estimates the bias more accurately than existing local explanation methods, and detects the increase/decrease in bias due to affirmative/punitive actions in the classifier.

\end{abstract}