### 2022-06-01 chapters/fairness/fif/fairxplainer.tex

\section{\fairXplainer: An Algorithm to Compute Fairness Influence Functions}\label{sec:fairxplainer}
\begin{comment}
Represent the algorithm as a sequence of two blocks: (1) Component function learning (2) Covariance computation of component functions. 
\end{comment}
We propose an algorithm, {\fairXplainer}, that leverages the variance decomposition of Eq.~\eqref{eq:fif_decompose} to compute the Fairness Influence Functions (FIFs) of all the subset of features. {\fairXplainer} has two algorithmic blocks: (i) local regression to decompose the classifier into component functions and (ii) computing the variance (or covariance) of each component. We describe the schematic of \fairXplainer{} in Algorithm~\ref{algo:framework}.
\setlength{\textfloatsep}{12pt}% Remove \textfloatsep

\noindent\textbf{A Set-additive Representation of the Classifier.} To apply variance decomposition (Eq.~\eqref{eq:variance_decomposition_set_notation}), we learn a set-additive representation of the classifier (Eq.~\eqref{eq:functional_decomposition_set_notation}). Let us denote the classifier $ \alg $ conditioned on a sensitive group $ \mathbf{a} $ as $ \function_{\mathbf{a}}(\nonsensitive) \triangleq \alg(\nonsensitive, \sensitive = \mathbf{a}) $. We express $ \function_{\mathbf{a}} $ as a set-additive model:
\begin{align}
\label{eq:set_additive_classifier}
\function_{\mathbf{a}}(\mathbf{X}) = \function_{\mathbf{a}, 0} +  \sum_{\mathbf{S} \subseteq [\numnonsensitive] \setminus \emptyset, |\mathbf{S}| \le \lambda} \function_{\mathbf{a},\mathbf{S}}(\mathbf{X}_{\mathbf{S}}) + \delta
\end{align}
\begin{algorithm}
	\caption{{\fairXplainer}}\label{algo:framework}
	\hspace*{\algorithmicindent} \textbf{Input:} $ \alg: (\nonsensitive, \sensitive) \rightarrow \widehat{Y},  \mathbf{D} = \{(\mathbf{x}^{(i)}, \mathbf{a}^{(i)}, y^{(i)})\}_{i=1}^n, f(\alg, \mathbf{D}) \in \mathbb{R}^{\geq 0}, \lambda $  \\
	\hspace*{\algorithmicindent} \textbf{Output:} $ w_{\mathbf{S}}$ 
	\begin{algorithmic}[1]
		\State $ \mathbf{a}_{\max} = \argmax_{\mathbf{a}} \Pr[\widehat{Y} = 1 | \sensitive = \mathbf{a}], \mathbf{a}_{\min} = \argmin_{\mathbf{a}} \Pr[\widehat{Y} = 1 | \sensitive = \mathbf{a}], k \gets    |\nonsensitive| $
		\label{algo_line:fif_computation_start}

		\For{$ \mathbf{a} \in \{\mathbf{a}_{\max}, \mathbf{a}_{\min}\} $} \Comment{Enumerate for specific sensitive groups}
		\State $ \function_{\mathbf{a}, \mathbf{S}},\function_{\mathbf{a}, 0} \gets \textsc{LocalRegression}(\alg(\nonsensitive, \sensitive = \mathbf{a}), \{\mathbf{x}^{(i)}\}_{i=1}^n, \lambda, k) $

		\State $ V_{\mathbf{a}, \mathbf{S}} \gets \textsc{Covariance}(\function_{\mathbf{a}},  \{\mathbf{x}^{(i)}\}_{i=1}^n, \function_{\mathbf{a}, \mathbf{S}},\function_{\mathbf{a}, 0}) $
		\EndFor
		\State Compute $ w_{\mathbf{S}}$ using  $V_{\mathbf{a}_{\max},\mathbf{S}}$ and $V_{\mathbf{a}_{\min},\mathbf{S}}$ as in Equation~\eqref{eq:fif_decompose}
		%\State $ w_{\mathbf{S}}  = (V_{\mathbf{a}_{\max},\mathbf{S}} - V_{\mathbf{a}_{\min},\mathbf{S}})/(1 - (\Pr[\widehat{Y} = 1 |  \sensitive = \mathbf{a}_{\max}] + \Pr[\widehat{Y} = 1 |  \sensitive = \mathbf{a}_{\min}])) $
		\label{algo_line:fif_computation_end}
		\Comment{Theorem~\ref{lemma:fif}}

%		\Statex
		\Function{LocalRegression}{$ g_\mathbf{a}, \{\mathbf{x}^{(i)}\}_{i=1}^n, \lambda, k $}
		\label{algo_line:local_regression_start}
		\State \textbf{Initialize}: $ \function_{\mathbf{a}, 0} \leftarrow \textsc{Mean}(\{g(\mathbf{x}^{(i)})\}_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n), \widehat{\function}_{\mathbf{a}, \mathbf{S}} \leftarrow 0, \forall \mathbf{S} \in [k]\setminus\emptyset, |\mathbf{S}|\le \lambda$ \label{alg_line:initialization}
		\While{each $ \widehat{\function}_{\mathbf{a},\mathbf{S}} $ does not converge}
		\For{each $\mathbf{S}$}
		\State $ \widehat{\function}_{\mathbf{a}, \mathbf{S}} \leftarrow \textsc{Smooth}[\{\function_{\mathbf{a}}(\mathbf{x}^{(i)}) - \function_{\mathbf{a}, 0} - \sum_{\mathbf{S} \ne \mathbf{S}'} \widehat{\function}_{\mathbf{a}, \mathbf{S}'}(\mathbf{x}_{\mathbf{S}}^{(i)})\}_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n] 
		$\label{alg_line:backfitting_step} \Comment{Backfitting}
		\State $\widehat{\function}_{\mathbf{a}, \mathbf{S}}  \leftarrow \widehat{\function}_{\mathbf{a}, \mathbf{S}} - \textsc{Mean}(\{\widehat{\function}_{\mathbf{a}, \mathbf{S}}(\mathbf{x}_{\mathbf{S}}^{(i)})\}_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n)$ \label{alg_line:mean_centered} \Comment{Mean centering}
		\EndFor
		\EndWhile
		\State \Return $ \function_{\mathbf{a}, \mathbf{S}},\function_{\mathbf{a}, 0} $
		\label{algo_line:local_regression_end} 
		\EndFunction

		\Function{Covariance}{$\function_{\mathbf{a}},  \{\mathbf{x}^{(i)}\}_{i=1}^n, \function_{\mathbf{a}, \mathbf{S}},\function_{\mathbf{a}, 0}$}
		\label{algo_line:covariance_computation_start}
		\State \Return $ \sum_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n \function_{\mathbf{a}, \mathbf{S}}(\mathbf{x}_{\mathbf{S}}^{(i)})(g_\mathbf{a}(\mathbf{x}^{(i)}) - \function_{\mathbf{a}, 0}) $
		\label{algo_line:covariance_computation_end}
		%		 \Comment{Eq.~\eqref{eq:covariance_computation}}
		\EndFunction
	\end{algorithmic}
\end{algorithm}
Here, $ \function_{\mathbf{a}, 0} $ is a constant and $ \function_{\mathbf{a},\mathbf{S}} $ is a \emph{component function} of $ \function_{\mathbf{a}} $ taking  $ \nonsensitive_{\mathbf{S}} $ as input, and $ \delta $ is the approximation error. For computational tractability, we consider only components of \emph{maximum order} $ \lambda $. {\fairXplainer} deploys backfitting algorithm for learning component functions in Eq.~\eqref{eq:set_additive_classifier}.

\noindent\textbf{Local Regression with Backfitting.} We perform local regression with backfitting algorithm to learn the component functions up to order $ \lambda $ (Line~\ref{algo_line:local_regression_start}--\ref{algo_line:local_regression_end}). Backfitting algorithm is an iterative algorithm, where in each iteration one component function, say $ \function_{\mathbf{a}, \mathbf{S}} $, is learned while keeping other component functions fixed. Specifically, $ \function_{\mathbf{a}, \mathbf{S}} $ is learned as a smoothed function of $ g $ and rest of the components $ \function_{\mathbf{a}, \mathbf{S}'} $, where $ \mathbf{S}' \ne \mathbf{S} $ is a non-empty subset of $ [k]\setminus \emptyset $. To keep every component function mean centered, backfitting requires to impose the constraints $ \function_{\mathbf{a}, 0} = \textsc{Mean}(\{g(\mathbf{x}^{(i)})\}_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n) $ (Line~\ref{alg_line:initialization}), which is the mean of $ \function_{\mathbf{a}} $ evaluated on samples belonging to the sensitive group $ \sensitive = \mathbf{a} $;  and $ \sum_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n\function_{\mathbf{a}, \mathbf{S}}(\mathbf{x}_{\mathbf{S}}^{(i)}) = 0$ (Line~\ref{alg_line:mean_centered}), where $ \mathbf{x}_{\mathbf{S}}^{(i)} $ be the subset of feature values associated with feature indices $ \mathbf{S} $ for the $ i $-th sample $ \mathbf{x}^{(i)} $. These constraints assign the expectation of $ \function_{\mathbf{a}} $ on the constant term $ \function_{\mathbf{a}, 0} $ and the variance of $ \function_{\mathbf{a}} $ to the component functions.

While performing local regression, backfitting uses a smoothing operator~\cite{loader2012smoothing} over the set of samples (Line~\ref{alg_line:backfitting_step}). A smoothing operator, referred as $ \textsc{Smooth} $, allows us to learn a global representation of a component function by smoothly interpolating the local curves obtained by local regression~\cite{loader2012smoothing}. In this paper, we apply cubic spline smoothing~\cite{li2010global} to learn each component function. Cubic spline is a piecewise polynomial of degree $ 3 $ with $ C^2 $ continuity. Hence, up to the second derivatives of each piecewise term are zero at the endpoints of intervals. We refer to Appendix~\ref{sec:smoothing} for further details.%\todo{Add pseudo-code of {\fairXplainer}}

\begin{comment}
\begin{algorithm}[!t]
	\caption{Local Regression with Backfitting Algorithm}
	\label{algo:backfitting_algorithm}
%	\hspace*{\algorithmicindent} \textbf{Input:} $ g_\mathbf{a}(\mathbf{X}), \{\mathbf{x}^{(i)}\}_{i=1}^n, \lambda $ \\
%	\hspace*{\algorithmicindent} \textbf{Output:} $ \function_{\mathbf{a}, 0}, \function_{\mathbf{a}, \mathbf{S}} $ 
	\begin{algorithmic}[1]
		\Procedure{LocalRegression}{$ g_\mathbf{a}, \{\mathbf{x}^{(i)}\}_{i=1}^n, \lambda, k $}
		\State \textbf{Initialize}: $ \function_{\mathbf{a}, 0} \leftarrow \frac{1}{n}\sum_{i=1}^n\function_{\mathbf{a}}(\mathbf{x}^{(i)}), \widehat{\function}_{\mathbf{a}, \mathbf{S}} \leftarrow 0, \forall \mathbf{S} \in [k]\setminus\emptyset, |\mathbf{S}|\le \lambda$ \label{alg_line:initialization}
		\While{each $ \widehat{\function}_{\mathbf{a},\mathbf{S}} $ does not converge}
		\For{each $\mathbf{S}$}
		\State $ \widehat{\function}_{\mathbf{a}, \mathbf{S}} \leftarrow \mathsf{Smooth}[\{\function_{\mathbf{a}}(\mathbf{x}^{(i)}) - \function_{\mathbf{a}, 0} - \sum_{\mathbf{S} \ne \mathbf{S}'} \widehat{\function}_{\mathbf{a}, \mathbf{S}'}(\mathbf{x}_{\mathbf{S}}^{(i)})\}_{i=1}^n] $\label{alg_line:backfitting_step} \Comment{Backfitting step}
		\State $\widehat{\function}_{\mathbf{a}, \mathbf{S}}  \leftarrow \widehat{\function}_{\mathbf{a}, \mathbf{S}} - \frac{1}{n}\sum_{i=1}^n \widehat{\function}_{\mathbf{a}, \mathbf{S}}(\mathbf{x}_{\mathbf{S}}^{(i)})$ \label{alg_line:mean_centered} \Comment{Mean centering}
		\EndFor
		\EndWhile
		\State \Return $ \function_{\mathbf{a}, 0}, \function_{\mathbf{a}, \mathbf{S}} $ 
		\EndProcedure
	\end{algorithmic}
\end{algorithm}
\end{comment}

\noindent\textbf{Variance and Covariance Computation.} Once each component function $ \function_{\mathbf{a}, \mathbf{S}} $ is learned with $ \textsc{LocalRegression} $ (Line~\ref{algo_line:local_regression_start}--\ref{algo_line:local_regression_end}), we compute variances of the component functions and their covariances with $ \function_{\mathbf{a}} $. Since each component function is mean centered (Line~\ref{alg_line:mean_centered}), we compute the variance of $ \function_{\mathbf{a}, \mathbf{S}} $ on the dataset as
$  \mathsf{Var}[\function_{\mathbf{a}, \mathbf{S}}] = \sum_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n (\function_{\mathbf{a}, \mathbf{S}}(\mathbf{x}_{\mathbf{S}}^{(i)}))^2 $. Hence, variance captures the independent effect of $ \function_{\mathbf{a}, \mathbf{S}} $.
%$ V_{\mathbf{a}, \mathbf{S}} $ is the variance of $ \function_{\mathbf{a}, \mathbf{S}} $ evaluated on each $ \mathbf{x}_{\mathbf{S}}^{(i)} $ in the dataset. 
Covariance is computed to account for the correlation among features $ \nonsensitive $. We compute the covariance of $ \function_{\mathbf{a}, \mathbf{S}} $ with $ \function_{\mathbf{a}} $ on the dataset as
$
\label{eq:covariance_computation}
	 \mathsf{Cov}[\function_{\mathbf{a}, \mathbf{S}}, \function_{\mathbf{a}}] = \sum_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n \function_{\mathbf{a}, \mathbf{S}}(\mathbf{x}_{\mathbf{S}}^{(i)})(g_\mathbf{a}(\mathbf{x}^{(i)}) - \function_{\mathbf{a}, 0}). 
$
Here, $ g_\mathbf{a}(\cdot) - \function_{\mathbf{a}, 0} $ is the mean centered form of $ \function_{\mathbf{a}} $. Covariance of $ \function_{\mathbf{a}, \mathbf{S}} $ can be both positive and negative depending on whether the features $ \nonsensitive_{\mathbf{S}} $ are positively or negatively correlated with $ \function_{\mathbf{a}} $. Specifically, under the set additive model, we obtain $ \mathsf{Cov}[\function_{\mathbf{a}, \mathbf{S}}, \function_{\mathbf{a}}] = \mathsf{Var}[\function_{\mathbf{a}, \mathbf{S}}] + \mathsf{Cov}[\function_{\mathbf{a}, \mathbf{S}}, \sum_{\mathbf{S} \ne \mathbf{S}'} \function_{\mathbf{a}, \mathbf{S}'}] $. Now, we use $ V_{\mathbf{a}, \mathbf{S}} =  \mathsf{Cov}[\function_{\mathbf{a}, \mathbf{S}}, \function_{\mathbf{a}}] $ as the effective variance of $ \nonsensitive_{\mathbf{S}} $ for a given sensitive group $ \sensitive = \mathbf{a} $ (Line~\ref{algo_line:covariance_computation_start}--\ref{algo_line:covariance_computation_end}). In Line~\ref{algo_line:fif_computation_start}--\ref{algo_line:fif_computation_end}, we compute $ V_{\mathbf{a}, \mathbf{S}} $ for the most and the least favored groups, and plug them in Theorem~\ref{lemma:fif} to compute FIF of $ \nonsensitive_{\mathbf{S}} $.

%In both variance and covariance decomposition, $ \mathsf{Var}[\function_{\mathbf{a}}] = \sum_{\mathbf{S} \subseteq [\numnonsensitive] \setminus \emptyset, |\mathbf{S}| \le \lambda} V_{\mathbf{a}, \mathbf{S}}$. The difference between variance and covariance decomposition is that $ V_{\mathbf{a}, \mathbf{S}} \in \mathbb{R}^{\ge 0} $ is positive in  variance decomposition, while $ V_{\mathbf{a}, \mathbf{S}} \in \mathbb{R} $ may also be negative in covariance decomposition. In covariance decomposition, we can separate between the structural and correlative effects of $ \function_{\mathbf{a}, \mathbf{S}} $. Formally, $ \mathsf{Cov}[\function_{\mathbf{a}, \mathbf{S}}, \function_{\mathbf{a}}] = \mathsf{Var}[\function_{\mathbf{a}, \mathbf{S}}] + \mathsf{Cov}[\function_{\mathbf{a}, \mathbf{S}}, \sum_{\mathbf{S}' \subseteq [\numnonsensitive] \setminus \emptyset, |\mathbf{S}'| \le \lambda, \mathbf{S} \ne \mathbf{S}'} \function_{\mathbf{a}, \mathbf{S}'}] $, where  to the right side of equality, first term refers to the structural effect and the second term refers to correlative effect of $ \function_{\mathbf{a}, \mathbf{S}} $.

\noindent\textbf{Extension of {\fairXplainer} to Equalized Odds and Predictive Parity.} 
We extend {\fairXplainer} in computing FIF of group fairness metrics beyond statistical parity, namely equalized odds and predictive parity. For equalized odds, we deploy {\fairXplainer} twice, one for computing FIFs on a subset of samples in the dataset where $ Y = 1 $ and another on samples with $ Y = 0 $. Then, the maximum of the sum of FIFs between $ Y = 1 $ and $ Y = 0 $ quantifies the equalized odds of the classifer.

To compute FIFs for predictive parity, we condition the dataset by the predicted class $ \widehat{Y} $ and separate into two sub-datasets: $ \widehat{Y} = 1 $ and  $ \widehat{Y} = 0 $. For each sub-dataset, we deploy $ \fairXplainer $ by setting the ground-truth class $ Y $ as label. This contrasts the computation for  statistical parity and equalized odds, where the predicted class $ \widehat{Y} $ is considered as label. Finally, the maximum of the sum of FIFs between two sub-datasets for $ \widehat{Y} = 1 $ and  $ \widehat{Y} = 0 $ quantifies the predictive parity of the classifier.