### 2022-06-01 chapters/fairness_background.tex
upto preliminaries of part 1

\part{Fairness in Machine Learning}
\chapter{Preliminaries}