### 2022-06-02 chapters/list_of_papers.tex

%\clearpage
%\begin{center}
%	\large\textbf{List of Publications}
%	\normalsize
%\end{center}

\section*{List of Publications}
\addcontentsline{toc}{chapter}{List of Publications}
This thesis is based on the following publications.

\begin{enumerate}

	\item ``How Biased is Your Feature?": Computing Fairness Influence Functions with Global Sensitivity Analysis\\
	\underline{Bishwamittra Ghosh}, Debabrota Basu, Kuldeep S. Meel\\
	Under review, 2022

	\item \href{https://arxiv.org/pdf/2205.06936.pdf}{Efficient Learning of Interpretable Classification Rules} \\
	\underline{Bishwamittra Ghosh}, Dmitry Malioutov, Kuldeep S. Meel\\
	In Proceedings of JAIR, 2022

	\item \href{https://arxiv.org/pdf/2109.09447.pdf}{Algorithmic Fairness Verification with Graphical Models} \\
	\underline{Bishwamittra Ghosh}, Debabrota Basu, Kuldeep S. Meel\\
	In Proceedings of AAAI, 2022

	\item \href{https://arxiv.org/pdf/2009.06516.pdf}{Justicia: A Stochastic SAT Approach to Formally Verify Fairness} \\
	\underline{Bishwamittra Ghosh}, Debabrota Basu, Kuldeep S. Meel\\
	In Proceedings of AAAI, 2021

	\item \href{https://bishwamittra.github.io/publication/ecai_2020/paper.pdf}{Classification Rules in Relaxed Logical Form
	} \\
	\underline{Bishwamittra Ghosh}, Dmitry Malioutov, Kuldeep S. Meel\\
	In Proceedings of ECAI, 2020

	\item \href{https://bishwamittra.github.io/publication/imli-ghosh.pdf}{IMLI: An Incremental Framework for MaxSAT-Based Learning of Interpretable Classification Rules}\\
	\underline{Bishwamittra Ghosh}, Kuldeep S. Meel\\
	In Proceedings of AIES, 2019

\end{enumerate}
\clearpage