### 2022-06-02 chapters/fairness/fvgm/applications.tex
tables and figures updated with short caption

\section{Applications of {\fvgm}}\label{fvgm_sec:applications}
In this section, we apply {\fvgm} for verifying fairness-enhancing algorithms and depreciation attacks. We also demonstrate that {\fvgm} facilitates computation of fairness influence functions by enabling the detection of bias due to individual features.

\iffalse
\begin{figure}[t!]
	%\vspace{-1em}

	\centering
	\subfloat{\includegraphics[scale=0.25]{figures/fairness/fvgm/disp_fairness_attack}}
	\subfloat{\includegraphics[scale=0.25]{figures/fairness/fvgm/stat_fairness_attack}}
	%\vspace{-0.5em}
	\caption{Verifying poisoning attack against fairness using {\fvgm}. The red line denotes the safety margin of the ML model against the attack.}\label{fvgm_fig:fairness_attacks}
\end{figure}
\fi

\paragraph{Verifying Fairness-enhancing Algorithms.} We deploy {\fvgm} in verifying the effectiveness of fairness-enhancing algorithms designed to ameliorate bias. For example, fairness 
pre-processing algorithms can be validated by applying {\fvgm} on the unprocessed and the processed data separately and comparing different fairness metrics. In Table~\ref{fvgm_tab:fair_algo_verification}, we report the effect of fairness algorithms w.r.t. four fairness metrics: disparate impact (DI), path-specific causal fairness (PCF), statistical parity (SP), and equalized odds (EO). Note that, fairness is improved if DI increases and the rest of the metrics decrease. For instance, in most instances for Adult dataset, reweighing (RW)~\cite{kamiran2012data} and optimized pre-processing (OP)~\cite{calmon2017optimized} algorithms are successful in improving fairness. The exceptional case is the unfairness regarding the sensitive feature `sex', where OP algorithm fails in improving fairness metric EO. Thus, \textit{{\fvgm} verifies the enhancement and decrement in fairness by  fairness-enhancing algorithms. }

\begin{figure}
	\centering
	\subfloat{\includegraphics[scale=0.3]{figures/fairness/fvgm/disp_fairness_attack}}
	\subfloat{\includegraphics[scale=0.3]{figures/fairness/fvgm/stat_fairness_attack}}

	\caption[Verifying fairness attacks]{Verifying poisoning attack against fairness using {\fvgm}. The red line denotes the safety margin of the ML model against the attack.}\label{fvgm_fig:fairness_attacks}
\end{figure}

\paragraph{Verifying Fairness Attacks.}
We apply {\fvgm} in verifying a fairness poisoning-attack algorithm. This algorithm injects a small fraction of poisoned samples into the training data such that the classifier becomes relatively unfair~\cite{solans2020poisoning}. We apply this attack to add $\{1, 5, \dots, 160\}$ poisoned samples  and measure the corresponding disparate impact and statistical parity. In Figure~\ref{fvgm_fig:fairness_attacks},  {\fvgm} verifies that the disparate impact of the classifier decreases and statistical parity increases, i.e.\textit{ the classifier becomes more unfair},\textit{ as the number of poisoned samples increases}. Therefore, {\fvgm} shows the potential of being deployed in safety-critical applications to detect fairness attacks. For example, if we set $0.9$ as an acceptable threshold of disparate impact, {\fvgm} can raise an alarm once $160$ poisoned samples are added.

\iffalse
\begin{table}[!t]

	\centering
	\small
	\caption{Verification of fairness algorithms using {\fvgm}. Numbers in bold refer to fairness improvement. RW and OP refer to reweighing and optimized-preprocessing algorithms.  }\label{fvgm_tab:fair_algo_verification}
	\setlength{\tabcolsep}{.4em}

	\begin{tabular}{lllrrrrrrrrrrrrr}

		\toprule
		Dataset & Sensitive & Algo. & $ \Delta $ DI &  $ \Delta $ PCF & $ \Delta $ SP & $ \Delta $ EO\\
		\midrule

		\multirow{4}{*}{Adult}&\multirow{2}{*}{race}&RW&$ \textbf{0.53} $&$ \textbf{-0.06} $&$ \textbf{-0.06} $&$ \textbf{-0.02} $\\
		&&OP&$ \textbf{0.57} $&$ \textbf{-0.07} $&$ \textbf{-0.07} $&$ \textbf{-0.02} $\\
		\cmidrule{2-7}
		&\multirow{2}{*}{sex}&RW&$ \textbf{0.96} $&$ \textbf{-0.16} $&$ \textbf{-0.15} $&$ \textbf{-0.08} $\\
		&&OP&$ \textbf{0.43} $&$ \textbf{-0.08} $&$ \textbf{-0.08} $&$ 0.03 $\\

		\midrule
		\multirow{4}{*}{COMPAS}&\multirow{2}{*}{race}&RW&$ \textbf{0.13} $&$ \textbf{-0.07} $&$ \textbf{-0.07} $&$ \textbf{-0.06} $\\
		&&OP&$ \textbf{0.15} $&$ \textbf{-0.08} $&$ \textbf{-0.08} $&$ \textbf{-0.05} $\\
		\cmidrule{2-7}
		&\multirow{2}{*}{sex}&RW&$ \textbf{0.1} $&$ \textbf{-0.04} $&$ \textbf{-0.04} $&$ 0.04 $\\
		&&OP&$ \textbf{0.09} $&$ \textbf{-0.04} $&$ \textbf{-0.04} $&$ \textbf{-0.03} $\\

		\midrule
		\multirow{4}{*}{German}&\multirow{2}{*}{age}&RW&$ \textbf{0.52} $&$ \textbf{-0.53} $&$ \textbf{-0.52} $&$ \textbf{-0.47} $\\
		&&OP&$ \textbf{0.53} $&$ \textbf{-0.53} $&$ \textbf{-0.53} $&$ \textbf{-0.51} $\\
		\cmidrule{2-7}
		&\multirow{2}{*}{sex}&RW&$ -0.06 $&$ 0.06 $&$ 0.06 $&$ 0.02 $\\
		&&OP&$ -0.12 $&$ 0.12 $&$ 0.12 $&$ 0.07 $\\

		\bottomrule
	\end{tabular}
\end{table}
\fi

\paragraph{Fairness Influence Function (FIF): Tracing Sources of Unfairness.}
Another application of \fvgm{} as a fairness verifier is to quantify the effect of a subset of features on fairness. Thus, we define fairness influence function (FIF)  that computes the effect of a subset of non-sensitive features $\mathbf{S} \subseteq \nonsensitive$ on the probability of positive prediction of a classifier given a specific sensitive group $ \sensitive = \mathbf{a}$,
$
	\mathsf{FIF}(\mathbf{S}) \triangleq \Pr[\hat{Y} = 1 | \sensitive = \mathbf{a}, \mathcal{D}] - \Pr[\hat{Y} = 1 | \sensitive = \mathbf{a},  \mathcal{D}_{-\mathbf{S}}].
$
FIF allows us to explain the sources of unfairness in the classifier. In practice, we compute FIF of $\mathbf{S}$ by replacing the probability distribution of $\mathbf{S}$ with a uniformly random distribution, referred to as  $ \mathcal{D}_{-\mathbf{S}} $, and reporting differences in the conditional probability of positive prediction of the classifier.

In Figure~\ref{fvgm_fig:influence function}, we compute FIF for all features in COMPAS dataset, separately for two sensitive groups: Female (`sex' $ = 1 $) and Male. This dataset concerns the likelihood of a person of re-offending crimes within the next two years. We first observe that the base values of the probability of positive prediction are different for the two groups ($ 0.46 $ vs $ 0.61 $ for Female and Male), thereby showing Male as more probable to re-offend crimes than Female. Moreover, FIF of the feature `age' is comparatively higher in magnitude for Male than Female. 
This implies that while deciding recidivism the algorithm assumes that Female individuals across different ages re-offend crimes with almost the same probability and the probability of re-offending for Male individuals highly depends on age. Thus, applying {\fvgm} to compute FIF provides us insights about sources of bias and thus, indicators to improve fairness.
$\Pr[\hat{Y} = 1|$age\_0 $\ge$ 0.5]

\begin{figure}[t!]	
	\centering
	\vspace{-2ex}
	\subfloat{\includegraphics[scale=0.3]{figures/fairness/fvgm/dependency_exp_Learn-dependency_compas_lr_sex_1}}
	\subfloat{\includegraphics[scale=0.3]{figures/fairness/fvgm/dependency_exp_Learn-dependency_compas_lr_sex_0}}
	%\vspace{-0.5em}
%	\vspace{-2ex}
	\caption[FIF computation using {\fvgm}]{FIF computation for COMPAS dataset. For Female (left plot) and Male, influence of `age' decreases and the probability of positive prediction of the classifier increases  by different magnitudes.}\label{fvgm_fig:influence function}
%	\vspace{-2ex}
\end{figure}