### 2022-06-07 chapters/fairness/justicia/table_synthetic.tex

\begin{table}
 \caption[Accuracy of {\justicia}]{Results on synthetic benchmark.  `\textemdash'~ refers that the verifier cannot compute the metric. The number in bold denotes the best result, where the estimated fairness metric is closest to the exact value.}
 \label{fairness_justicia_tab:synthetic}
    \centering
%        \vspace*{-.2em}
%        \setlength{\tabcolsep}{.1em}
            \begin{tabular}{ccccccc}
                \toprule
                Metric & Exact  & {\justicia} & FairSquare & VeriFair & AIF360\\
                 \midrule
				Disparate impact &  $ 0.26 $  &  $ \textbf{0.25} $    &  $ 0.99 $  &  $ 0.99 $  &  $ \textbf{0.25} $  \\
				Statistical parity &  $ 0.53 $  &  $ \textbf{0.54} $    & \textemdash & \textemdash &  $ \textbf{0.54} $  \\
                \bottomrule
    \end{tabular}

\end{table}