### 2022-06-08 chapters/interpretability_background.tex

\part{Interpretable Machine Learning}

In this part of the thesis, we discuss preliminaries in interpretable machine learning in Chapter~\ref{chapter_interpretability_preliminaries}, an incremental learning framework for interpretable classification rules in Chapter~\ref{chapter:imli}, and extend incremental learning to more expressible classification rules in Chapter~\ref{chapter:crr}.

\chapter{Preliminaries}
\label{chapter_interpretability_preliminaries}
%\section{Preliminaries}
%\label{interpretability_imli_sec:preliminaries}

In this chapter, we discuss preliminaries for learning interpretable rule-based classifiers. Throughout this part of the thesis, we use capital  boldface letters such as $\mathbf{X}$ to denote matrices while lower boldface letters such as $\mathbf{x}$ are reserved for vectors/sets.

\subsection{Propositional Satisfiability}

Let $F$ be a Boolean formula defined over Boolean variables $\mathbf{b} = \{b_1,b_2,\dots ,b_m \}$. A literal $ v $ is a variable $b$ or its complement $\neg b$, and a clause $ C $ is a disjunction ($ \vee $) or a conjunction ($ \wedge $) of literals.\footnote{While \emph{term} is reserved to denote disjunction of literals in the literature, we use clause for both disjunction and conjunction.}  $F$ is in Conjunctive Normal Form (CNF) if $F \triangleq  \bigwedge_i C_i$ is a conjunction of clauses where each clause $C_i \triangleq  \bigvee_j v_j $  is a disjunction of literals. In contrast, $ F $ is in Disjunctive Normal Form (DNF) if $F \triangleq  \bigvee_i C_i$ is a disjunction of clauses where each clause $C_i \triangleq  \bigwedge_j v_j $ is a conjunction of literals. We use $\sigma$ to denote an assignment of variables in  $\mathbf{b}$ where $ \sigma(b_i) \in $ \{true, false\}. A \emph{satisfying assignment} $ \sigma^* $ of $F$ is an assignment  that evaluates $F$  to true and is denoted by $ \sigma^* \models F $. The propositional satisfiability (SAT) problem finds a satisfying assignment $ \sigma^* $ to a CNF formula $ F $  such that $ \forall i, \sigma^* \models C_i $, wherein $ \sigma^*\models C_i $ if and only if $ \exists v_j \in C_i, \sigma^*(v_j) = \text{true} $. Informally, $ \sigma^* $ satisfies at least one literal in each clause of a CNF.

\subsection{Relaxation of Logical Formulas}

A hard-OR or a soft-AND clause is a tuple $ (C,\eta) $ with an extra parameter $ \eta $, where $ \eta $ is the threshold on the literals in $ C $. Let $ \mathds{1}[\text{true}]=1 $ and $ \mathds{1}[\text{false}]=0 $. Motivated by Boolean cardinality constraints, we propose relaxed-CNF formulas, where in addition to $F$, we have two more parameters $\eta_c$ and $\eta_l$. We say that $(F,\eta_c, \eta_l)$ is in relaxed-CNF and $ \sigma^* $ is its satisfying assignment if and only if $\sigma^* \models (F,\eta_c, \eta_l)$ whenever $\sum_{i=1}^k \mathds{1}[\sigma^* \models (C_i,\eta_l)] \ge \eta_c$, where $\sigma^* \models (C_i, \eta_l)$ if and only if $\sum_{v \in C_i} \mathds{1}[\sigma^* \models v] \geq \eta_l$. Informally,  $\sigma^*$ satisfies a clause $(C_i, \eta_l)$ if  at least $\eta_l$ literals in $C_i$ are set to true by $\sigma^*$ and $\sigma^*$ satisfies $(F,\eta_c, \eta_l)$ if at least $\eta_c$  clauses out of all $\{(C_i,\eta_l)\}$ clauses are true.

\begin{theorem}[\cite{benhamou1994two}]
	\label{interpretability_crr_thm:succinctness}	
	Let $ (C,\eta) $ be a clause in relaxed-CNF where  $ C $ has $ m  $ literals and $ \eta \in \{1,\dots,m\} $ is the threshold on literals.  An equivalent compact encoding of $ (C,\eta) $  into a CNF formula $ F= \bigwedge_i C_i $ requires $ {m \choose{m-\eta+1}} $  clauses where each clause is distinct and has $ m-\eta+1 $ literals of $ (C,\eta) $. Therefore,  the total number of literals in $ F $ is $ (m-\eta+1){m\choose{m-\eta+1}} = m $ when $ \eta=1 $, otherwise $ (m-\eta+1){m\choose{m-\eta+1}} > m $.
\end{theorem}

\subsection{Inner Product}
In this part of the thesis, we use true as $ 1 $ and false as $ 0 $ interchangeably. Between two vectors  $\mathbf{u}$ and $\mathbf{v}$ of the same length defined over Boolean variables or constants (such as  $ 0 $ and $ 1 $), we define $\mathbf{u} \circ \mathbf{v} $  to refer to  their inner product. Formally, $\mathbf{u} \circ \mathbf{v} \triangleq \bigvee_{i} (u_{i} \wedge v_{i})$ is a disjunction of element-wise conjunction where $u_{i}$ and $v_{i}$ denote the $ i^\text{th} $ variable/constant of $\mathbf{u}$ and $\mathbf{v}$, respectively. In this context, the conjunction ``$\wedge$'' between a variable and a constant follows the standard interpretation: $b \wedge 0 = 0$ and $b \wedge 1 = b$.  For example, if $ \mathbf{u} = [u_1, u_2, u_3] $ and $ \mathbf{v} = [0,1,1] $, then $ \mathbf{u} \circ \mathbf{v} = u_2 \vee u_3 $. Applying numerical interpretation, the inner product can also be expressed as $\mathbf{u} \circ \mathbf{v} \triangleq \sum_{i} (u_{i} \cdot v_{i}) $. Hence, for the running example $ \mathbf{u} = [u_1, u_2, u_3] $ and $ \mathbf{v} = [0,1,1] $, we derive $ \mathbf{u} \circ \mathbf{v} = u_2  + u_3 $. In this part of the thesis, we use Boolean interpretation of the inner product for learning CNF classification rules in Chapter~\ref{chapter:imli} and numerical interpretation for learning relaxed-CNF classification rules in Chapter~\ref{chapter:crr}.

\subsection{MaxSAT}

The MaxSAT problem is an optimization analog to the SAT problem that finds an optimal assignment satisfying the maximum number of clauses in a CNF. In interpretable machine learning, we consider a \emph{weighted} variant of the MaxSAT problem\textemdash more specifically, a weighted-partial MaxSAT problem\textemdash that optimizes over a set of hard and soft constraints in the form of a weighted-CNF formula. In a weighted-CNF formula, a weight $ {\weight{C_i}} \in \mathbb{R}^+ \cup \{\infty\} $ is defined over each clause $ C_i $, wherein $C_i$ is called a \emph{hard} clause if $\weight{C_i} = \infty$, and  $C_i$ is  a \emph{soft} clause, otherwise.  To avoid notational clutter, we overload {\weight{\cdot}} to denote the weight of an assignment $ \sigma $. In particular, we define $\weight{\sigma}$ as the sum of the weights of \emph{unsatisfied clauses} in a CNF, formally, $$\weight{\sigma} = \sum_{i | \sigma \not \models C_i} \weight{C_i}.$$

Given a weighted-CNF formula $F \triangleq \bigvee_i C_i$ with weight $ \weight{C_i} $, the weighted-partial MaxSAT problems finds an optimal assignment $\sigma^*$ that achieves the minimum weight. Formally,  $\sigma^* = \MaxSAT(F,\weight{\cdot})$ if $\forall \sigma \neq \sigma^*, \weight{\sigma^*} \leq \weight{\sigma}$. The optimal weight of the MaxSAT problem $  \weight{\sigma^*} $ is infinity ($ \infty $) when at least one hard clause becomes unsatisfied.  Therefore,  the weighted-partial MaxSAT problem finds $ \sigma^* $ that satisfies all hard clauses\footnote{In our formulation, we assume that there is a satisfying assignment to a CNF formula containing all hard clauses.} and as many soft clauses as possible such that the total weight of unsatisfied soft clauses is minimum. In Chapter~\ref{chapter:imli}, we reduce the classification problem  to the solution of a weighted-partial MaxSAT problem. The knowledge of the inner workings of {MaxSAT} solvers is  not required for this chapter.

%
%\subsection{\red{Mixed Integer Linear Programming (MILP)}}
%\red{[Not required]}The MILP problem is an optimization problem operating on both integer and real-valued variables. Given an objective function and a set of constraints, a (minimization) MILP finds an optimal assignment of variables that minimizes the objective function.

\subsection{Rule-based Classification}

We consider a standard binary classification problem where $ \mathbf{X} \in \{0,1\}^{n\times m} $ is a binary feature matrix\footnote{Our framework allows learning on real-valued and categorical features, as discussed in Section~\ref{interpretability_imli_sec:non-binary}.} defined over Boolean features $\mathbf{x} = \{x_1, x_2, \dots, x_m\}$ and $ \mathbf{y} \in \{0,1\}^n $ is a vector of binary class labels.  The $ i^\text{th} $ row of $ \mathbf{X} $, denoted by $ \mathbf{X}_i \in \mathcal{X} $, is called a sample generated from a product distribution of all features $ \mathcal{X} $. Thus, $ \mathbf{X}_i $ is the valuation of features in $ \mathbf{x} $ and $ y_i $ is its class label.  $ \mathbf{X}_i $ is called a positive sample if $ y_i = 1 $, and a negative sample otherwise.

A classifier $ \Rule: \mathbf{x} \rightarrow \widehat{y} \in  \{0,1\} $ is a  function that takes a Boolean vector as input and outputs the predicted class label $\widehat{y}$.  %We define two rules  $ \R_1 $ and $ \R_2 $ to be  equivalent  if $ \forall i,\R_1(\X_i)=\R_2(\X_i) $. 
In a rule-based classifier, we view $ \mathcal{R} $ as a propositional formula defined over Boolean features $ \mathbf{x} $ such that, if $ \Rule $ becomes true for an input (or an assignment), the prediction $ \widehat{y} = 1 $ and $ \widehat{y} = 0 $, otherwise.  The goal is to design $\Rule$ not only to approximate the training data, but also to generalize to unseen samples arising from the same distribution. Additionally, we prefer to learn a \emph{sparse} $ \Rule $ in order to favor interpretability. We define the structural complexity of $ \Rule $, referred to as \emph{rule-size}, in terms of the number of literals it contains. Let $\clause(\Rule,i)$ denote  the  $i^\text{th}$ clause of $\Rule$ and $ |\clause(\Rule,i)| $ be  the number of literals in $\clause(\Rule,i)$. Thus, the size of the classifier is $|\Rule| =  \Sigma_i |\clause(\Rule,i)| $, which is the sum of literals in all clauses in $ \Rule $.

\begin{example}
	\normalfont
	Let $ \Rule \triangleq (x_1 \vee x_3) \wedge (\neg x_2 \vee x_3) $ be a CNF classifier defined over three Boolean features $ \{x_1, x_2, x_3\} $. For an input $ [0, 1, 1] $, the prediction of $ \Rule  $ is $ 1 $, whereas for an input $ [1, 1, 0] $, the prediction is $ 0 $ because $ \Rule $ is true and false in these two cases, respectively. Moreover, $ |\Rule| = 2 + 2 = 4 $ is the size of $ \Rule $. 
\end{example}

\paragraph{Decision Lists.} A decision list is a rule-based classifier consisting of ``if-then-else'' rules. Formally, a decision list~\cite{rivest1987learning} is an ordered list $ \Rule_L $ of pairs $ (C_1, v_1), \dots, (C_k, v_k) $ where $ C_i $ is a conjunction of literals (alternately, a single clause in a DNF formula) and $ v_i \in \{0,1\} $ is  a Boolean class label\footnote{In a more practical setting, $ v_i \in \{0, \dots, N\} $  can be multi-class for $ N \ge 1 $.}. Additionally, the last clause $ C_k \triangleq  $ true, thereby $ v_k $ is the default class. A decision list is defined as a classifier with the following interpretation: for a feature vector $ \mathbf{x} $,  $ \Rule_L(\mathbf{x}) $ is equal to $ v_i $ where $ i $ is the least index such that $ \mathbf{x} \models C_i $. Since the last clause is true, $ \Rule_L(\mathbf{x}) $ always exists. Intuitively, whichever clause (starting from the first) in $ \Rule_L $ is satisfied for an input, the associated class-label is considered as its prediction.

\paragraph{Decision sets.} A decision set is a set of \textit{independent} ``if-then'' rules. Formally,  a decision set $ \Rule_S $ is a set of pairs $ \{(C_1, v_1), \dots, (C_{k-1}, v_{k-1})\}  $ and a  default pair $ (C_k, v_k) $, where $ C_i $ is\textemdash similar to a decision list\textemdash a conjunction of literals and $ v_i \in \{0,1\} $ is a Boolean class label. In addition, the last clause $ C_k \triangleq  $ true and $ v_k $ is the default class. For a  decision set, if an input $ \mathbf{x} $ satisfies one clause, say $ C_i $, then the prediction is $ v_i $. If $ \mathbf{x} $ satisfies no clause, then the prediction is the default class $ v_k $. Finally, if $ \mathbf{x} $ satisfies  $ \ge 2 $ clauses, $ \mathbf{x} $ is assigned a class using a tie-breaking~\cite{lakkaraju2016interpretable}.

%\todo{End}