### 2022-07-15 chapters/interpretability/relaxed-cnf/conclusion.tex

\section{Chapter Summary}
\label{interpretability_crr_sec:conclusion}

In this chapter, we proposed an efficient combinatorial framework, called {\crr}, for learning relaxed-CNF  classification rules. Relaxed-CNF rules are  more expressive  than CNF/DNF rules.  {\crr}  uses a novel integration of mini-batch learning procedure with the MILP framework to learn sparse relaxed-CNF rules. Our experimental results demonstrate that {\crr} is able to learn relaxed-CNF  rules with higher accuracy and more concise representation than CNF rules. Moreover, the generated rules are sparser than decision lists in large datasets.