### 2023-02-07 chapters/fairness/fif/framework.tex

%\clearpage
\section{Fairness Influence Functions: Formulation and Properties}\label{fairness_fairXplainer_sec:fifs}
We formalize Fairness Influence Functions (FIF) as a quantifier of the contribution of a subset of features to the resultant bias of a classifier applied on a dataset.  In GSA, we observe that the variance of the output of a function can be attributed to the corresponding subset of input variables through variance decomposition (Eq.~\eqref{fairness_fairXplainer_eq:variance_decomposition_set_notation}). To leverage the power of GSA in fairness in machine learning, we first express the existing fairness metrics in terms of the variance of classifier's prediction in Section~\ref{fairness_fairXplainer_sec:metric_as_variance}. This allows us to formulate FIF by leveraging variance decomposition (Section~\ref{fairness_fairXplainer_sec:fif_formulation}).

\subsection{Fairness Metrics as the Variance of Prediction}
\label{fairness_fairXplainer_sec:metric_as_variance}

First, we observe that the random variable central to computing statistical parity is $ \mathds{1}[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}] $.
We refer to this indicator function as \textit{Conditional Positive Prediction (CPP)} of a classifier. Now, we express statistical parity as a functional of the probability of CPPs for different sensitive groups~\cite{ghosh2020justicia,benesse2021fairness}. For brevity, we defer similar formulations for the other fairness metrics, i.e. equalized odds and predictive parity, to Appendix~\ref{fairness_fairXplainer_app:pp}. %CPP has an additional conditional $ Y = y $.}
%For indicator functions, probability and expectation are equal.

%To this end, existing fairness metrics such as statistical parity is the first moment (probability or expectation) of prediction distribution while variance is the second moment of distribution.  
Our key idea for computing FIFs of features is to represent fairness metrics using the variance of CPPs. Formally, we express statistical parity using the variance of CPPs in Lemma~\ref{fairness_fairXplainer_lm:sp_var_relation}.

\begin{lemma}[Statistical Parity as Difference of Variances of CPPs]
	\label{fairness_fairXplainer_lm:sp_var_relation}
	Let $ \mathbf{a}_{\max} = \argmax_{\mathbf{a}} \Pr[\widehat{Y} = 1 \mid  \sensitive = \mathbf{a}] $ and $ \mathbf{a}_{\min} = \argmin_{\mathbf{a}} \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}] $ be the most and the least favored sensitive groups, respectively. The statistical parity of a binary\footnote{For a multi-class classifier, statistical parity of the target class $ y \in \mathbb{N} $ is $ \frac{\mathsf{Var}[\widehat{Y} = y \mid \sensitive = \mathbf{a}_{\max}]}{1 - \Pr[\widehat{Y} = y \mid  \sensitive = \mathbf{a}_{\max}]} - \frac{\mathsf{Var}[\widehat{Y} = y \mid \sensitive = \mathbf{a}_{\min}] }{1 - \Pr[\widehat{Y} = y \mid  \sensitive = \mathbf{a}_{\min}]}. $} classifier is the difference in the scaled variance of CPPs. Formally, if $\Pr[\widehat{Y} = 0 \mid  \sensitive = \mathbf{a}_{\max}] \neq 0$ and $\Pr[\widehat{Y} = 0 \mid  \sensitive = \mathbf{a}_{\min}] \neq 0$,
	\[
	 f_{\mathsf{SP}}(\alg, \mathbf{D}) = \frac{\mathsf{Var}[\widehat{Y} = 1\mid\sensitive = \mathbf{a}_{\max}]}{\Pr[\widehat{Y} = 0 \mid  \sensitive = \mathbf{a}_{\max}]} - \frac{\mathsf{Var}[\widehat{Y} = 1\mid\sensitive = \mathbf{a}_{\min}] }{\Pr[\widehat{Y} = 0 \mid  \sensitive = \mathbf{a}_{\min}]}.
	\]
\end{lemma}
\begin{proof}
	For a sensitive group $ \mathbf{a} $, CPP  is a Bernoulli random variable, where probability\footnote{For any binary event $ E $, expectation and probability are identical, $ \mathbb{E}[\mathds{1}(E)] = \Pr[E = 1] $.} $ p_{\mathbf{a}}  = \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}] $ and variance $ V_{\mathbf{a}} = \mathsf{Var}[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}] = p_{\mathbf{a}} (1 - p_{\mathbf{a}}) $. Thus, for sensitive groups $ \mathbf{a}_{\max} $ and $ \mathbf{a}_{\min} $, the statistical parity of the classifier is $ p_{\mathbf{a}_{\max}}  - p_{\mathbf{a}_{\min}} =  \frac{V_{\mathbf{a}_{\max}}}{1 - p_{\mathbf{a}_{\max}}}  - \frac{V_{\mathbf{a}_{\min}}}{1 - p_{\mathbf{a}_{\min}}} $. Replacing $ 1 - p_\mathbf{a} = \Pr[\widehat{Y} = 0 \mid \sensitive = \mathbf{a}] $  proves the lemma.
\end{proof}

\begin{repexample}{fairness_fairXplainer_ex:motivating_example}[Revisited]
	From Figure~\ref{fairness_fairXplainer_fig:dt_original}, the probability of CPPs of the decision tree for the most and least favored groups are $  \Pr[\widehat{Y} = 1 \mid \text{age = young}] = 0.695 $ and $ \Pr[\widehat{Y} = 1 \mid \text{age = elderly}] = 0.165 $, respectively. Thus, the statistical parity is $ 0.695 - 0.165 =  0.53 $. Next, we compute variance of CPPs as $  \mathsf{Var}[\widehat{Y} = 1\mid \text{age = young}] = 0.212 $ and $  \mathsf{Var}[\widehat{Y} = 1\mid \text{age = elderly}] = 0.138 $. Thus, following Lemma~\ref{fairness_fairXplainer_lm:sp_var_relation}, we compute the difference in the scaled  variance of CPPs as $ \frac{0.212}{1 - 0.695} - \frac{0.138}{1 - 0.165} =  0.529 \approx 0.53 $, which coincides with the statistical parity reported earlier.
\end{repexample}

\subsection{Formulation of FIF}
\label{fairness_fairXplainer_sec:fif_formulation}
We are given a binary classifier  $\alg $, a dataset $ \mathbf{D} $, and a  fairness metric $ f(\alg, \mathbf{D}) \in \mathbb{R}^{\geq 0} $. Our objective is to compute the influences of  features on $ f $. Particularly, we compute the influence of each subset of  features $ \feature_{\mathbf{S}} $, where $ \mathbf{S} = \{S_i \mid 1 \le |S_i| \le \numfeatures\} \subseteq [\numfeatures]  $ is a non-empty subset of indices of $ \feature $.

\begin{definition} \textbf{Fairness Influence Function (FIF)}\footnote{For equalized odds, a precise definition of FIF is $ w_{\mathbf{S}}: (\alg, Y, \feature_{\mathbf{S}}) \rightarrow \mathbb{R} $ taking the ground class $ Y $ as an additional input. For predictive parity, FIF is defined as $ w_{\mathbf{S}}: (\alg, \widehat{Y}, \feature_{\mathbf{S}}) \rightarrow \mathbb{R} $ taking the predicted class $ \widehat{Y} $ as an additional input.}, denoted by $ w_{\mathbf{S}}: (\alg, \feature_{\mathbf{S}}) \rightarrow \mathbb{R} $, measures the quantitative contribution of features $ \feature_{\mathbf{S}} \subseteq \feature $ on the incurred bias $ f(\alg, \mathbf{D}) $. Leveraging the variance-difference representation of $ f(\alg, \mathbf{D}) $ (Lemma~\ref{fairness_fairXplainer_lm:sp_var_relation}) and variance decomposition (Eq.~\eqref{fairness_fairXplainer_eq:variance_decomposition_set_notation}), we particularly define $ w_{\mathbf{S}} $ as
\begin{equation}\label{fairness_fairXplainer_eq:fif_decompose}
	w_{\mathbf{S}}  \triangleq \frac{V_{\mathbf{a}_{\max}, \mathbf{S}}}{ \Pr[\widehat{Y} = 0 \mid  \sensitive = \mathbf{a}_{\max}]} - \frac{V_{\mathbf{a}_{\min}, \mathbf{S}} }{\Pr[\widehat{Y} = 0 \mid  \sensitive = \mathbf{a}_{\min}]}.
\end{equation}
Here, given a sensitive group $ \mathbf{a} $, $ V_{\mathbf{a}, \mathbf{S}} \triangleq  \mathsf{Var} _{\feature_{\mathbf{S}}}[\mathbb{E}_{\feature\setminus\feature_{\mathbf{S}}}[\widehat{Y} = 1\mid \sensitive = \mathbf{a}, \feature_{\mathbf{S}}]]- \sum_{\mathbf{S}' \subset \mathbf{S}}{V} _{\mathbf{a}, \mathbf{S}'} $ is the contribution (decomposed variance) of features $ \feature_{\mathbf{S}} $ in $ \mathsf{Var}[\widehat{Y} = 1\mid\sensitive = \mathbf{a}] $.
\end{definition}

Informally, the FIF $ w_{\mathbf{S}} $ of $ \feature_{\mathbf{S}} $ is the difference in the scaled decomposed variance of CPPs between sensitive groups $ \mathbf{a}_{\max} $ and $ \mathbf{a}_{\min} $ as induced by  $ \feature_{\mathbf{S}} $. Thus, FIF of features is non-zero when the scaled decomposed variance-difference of CPPs is non-zero for those features, and vice versa. We refer to $ w_{\mathbf{S}} $ as an \emph{individual influence} when $ |\mathbf{S}| = 1 $, and as an \emph{intersectional influence} when $ |\mathbf{S}| > 1 $. Being able to naturally quantify the higher-order influences allow FIFs to explain the sources of bias in a more fine-grained manner. We experimentally validate this in Section~\ref{fairness_fairXplainer_sec:experiments}.

\paragraph{Properties of FIF} FIF as defined in Eq.~\eqref{fairness_fairXplainer_eq:fif_decompose} yields interesting properties, such as decomposability\footnote{Decomposability property is also known as the efficiency property in the context of Shapley values~\cite{roth1988shapley}.}, symmetry, and null properties, which we formally state in Theorem~\ref{fairness_fairXplainer_thm:fif_property}.

\begin{theorem}[Properties of FIF]
	\label{fairness_fairXplainer_thm:fif_property}
	Let $ f(\alg, \mathbf{D}) $ be the bias/unfairness of the classifier $ \alg $ on dataset $ \mathbf{D} $ according to linear group fairness metrics such as statistical parity. Let $ w_{\mathbf{S}}  $ be the FIF of a subset of  features $ \feature_{\mathbf{S}} $ as defined in Eq.~\eqref{fairness_fairXplainer_eq:fif_decompose}. 
	\begin{enumerate}
		\item[(a)] \textit{The decomposability property} of FIF states that the sum of FIFs of all subset of  features is equal to the bias of the classifier. 
		\begin{align}\label{fairness_fairXplainer_eq:constraint_framework}
		\sum_{\mathbf{S} \subseteq [\numfeatures] } w_{\mathbf{S}} = f(\alg, \mathbf{D})
		\end{align}
		\item[(b)] \textit{The symmetry property} states that  two features $ Z_i $ and $ Z_j $ are equivalent based on FIF if the sum of corresponding individual influences and the intersectional influences with all other features are the same. Mathematically,
		\begin{align}
		\sum_{\mathbf{S}'' \subseteq [m]\setminus\{i,j\}}w_{\mathbf{S}''\cup \{i\}} = \sum_{\mathbf{S}'' \subseteq [m]\setminus\{i,j\}}w_{\mathbf{S}''\cup \{j\}}  
		\end{align}
		if $\sum_{\mathbf{S}' \subseteq \mathbf{S}\cup \{i\}} w_{\mathbf{S}'} = 		\sum_{\mathbf{S}' \subseteq \mathbf{S} \cup \{j\}} w_{\mathbf{S}'}$ for every non-empty subset $ \mathbf{S} $ of $ [\numfeatures] $ containing neither $ i $ nor $ j $. 
 		\item[(c)] \textit{The null property} of FIF states that feature $ X_i $ s a dummy or neutral feature if sum of its individual influence and the intersectional influences with all other features is zero. Mathematically,
	\begin{align}
	 \sum_{\mathbf{S}'' \subseteq [m]\setminus \{i\}}w_{\mathbf{S}''\cup \{i\}} = 0 	 
	\end{align}
	if	$\sum_{\mathbf{S}' \subseteq \mathbf{S} \cup \{i\}} w_{\mathbf{S}'} = 		\sum_{\mathbf{S}' \subseteq \mathbf{S}} w_{\mathbf{S}'}$ for every non-empty subset $ \mathbf{S} $ of $ [\numfeatures] $ that does not contain $ i $.
\end{enumerate}
\end{theorem}
We emphasize that the decomposability property proposed here is global, i.e. it holds for a whole dataset, but the one used for Shapley-value based explanations~(ref. Def. 1, \cite{lundberg2017unified}), or any local explanation method~\cite{han2022explanation}, is specific to a given data point~\cite{sliwinski2019axiomatic}. Thus, they are fundamentally different.

The symmetry and null properties of FIFs also distinguish the FIF quantification proposed in Eq.~\eqref{fairness_fairXplainer_eq:fif_decompose} from the attribution methods considering only individual features, like SHAP. For example, a feature $i$ has zero impact on bias if not only its individual influence but the sum of influences of all the subsets of features that includes $i$ is zero. Thus, the symmetry and null properties stated here are by default global and intersectional, and these two aspects are absent in the existing bias explainers.

\paragraph{Bias Amplifying and Eliminating Features.} The sign of $ w_{\mathbf{S}} $ indicates whether features $ \feature_{\mathbf{S}} $ amplify the bias of the classifier or eliminate it. When the scaled decomposed variance of CPPs w.r.t.\ features $ \feature_{\mathbf{S}} $ for the sensitive group $ \mathbf{a}_{\max} $ is higher than the group $ \mathbf{a}_{\min} $, $ w_{\mathbf{S}} > 0 $. As such, $ \feature_{\mathbf{S}} $ increase bias. Conversely, when $ w_{\mathbf{S}} < 0 $, $ \feature_{\mathbf{S}} $ eliminates bias, and improves fairness. Finally, features $ \feature_{\mathbf{S}} $ are neutral in bias when $ w_{\mathbf{S}} = 0 $.

Now, we present the following two propositions that relate the sign of $ w_{\mathbf{S}} $ with the decomposed variance of CPPs.

\begin{proposition}\label{fairness_fairXplainer_prop:neg_fif}
	When $ w_{\mathbf{S}} < 0 $, i.e. features $ \feature_{\mathbf{S}} $ decrease bias, the decomposed variance of CPPs w.r.t.\ $ \feature_{\mathbf{S}} $ follows $ V_{\mathbf{a}_{\max}, \mathbf{S}} < V_{\mathbf{a}_{\min}, \mathbf{S}}  $.
\end{proposition}
Proposition~\ref{fairness_fairXplainer_prop:neg_fif} implies that if a subset of features $ \feature_{\mathbf{S}} $ is bias-eliminating, the conditional variance in positive outcomes induced by $ \feature_{\mathbf{S}} $ is smaller for the most favoured group than that of the least favoured group.
\begin{proposition}\label{fairness_fairXplainer_prop:pos_fif}
{If the decomposed variance of CPPs w.r.t.\ $ \feature_{\mathbf{S}} $ satisfies $ V_{\mathbf{a}_{\max}, \mathbf{S}} > {V_{\mathbf{a}_{\min}, \mathbf{S}}} $, the corresponding FIF $ w_{\mathbf{S}} > 0  $, i.e. features $ \feature_{\mathbf{S}} $ increase bias.}
\end{proposition}
Proposition~\ref{fairness_fairXplainer_prop:pos_fif} implies that if the conditional variance in positive outcomes induced by $ \feature_{\mathbf{S}} $ is larger for the most favoured group than that of the least favoured group, then this subset of features $ \feature_{\mathbf{S}} $ is bias-inducing.

\paragraph{Special Cases.} Since our FIF formulation is based on the variance of prediction, for (degenerate) cases when the conditional prediction of the classifier is always positive or always negative for any sensitive group, the variance of prediction becomes zero. This observation results in following two propositions.

\begin{proposition}[Perfectly Unbiased Classifiers]
	When $ \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}_{\max} ] = \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}_{\min} ] $ and both conditional probabilities are either $ 0 $ or $ 1 $, the FIF $w_{\mathbf{S}}  = 0 $ for all subsets of features $ \feature_{\mathbf{S}} $.
\end{proposition}
When $ \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}_{\max} ] = \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}_{\min} ] $ for a classifier, it means that the classifier equally yields positive predictions for each of the sensitive groups. Thus, there is no bias (in terms of statistical parity) in the classifier outcome. In that case, our formulation of FIF yields $w_{\mathbf{S}}  = 0 $ for all the subsets of features, and leads to a degenerate conclusion that all subsets of features are neutral or zero-bias inducing.
\begin{proposition}[Perfectly Biased Classifier]
	{When statistical parity is $ 1 $, i.e. $ \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}_{\max} ] = 1 $ and $ \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}_{\min} ] = 0 $, the sensitive features $ \sensitive $ are solely responsible for bias. In this case, the FIF $ w_{\mathbf{S}} = 0 $ for all features.}
\end{proposition}
When a classifier always yields positive predictions for the most favoured group and only negative predictions for the least favoured group, we obtain $ \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}_{\max} ] = 1 $ and $ \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}_{\min} ] = 0 $. This is a perfectly biased classifier, which can be expressed by the binary rule $\widehat{Y} = 1$ if $\sensitive = \mathbf{a}_{\max}$ and $0$ otherwise. As the discrimination is only due to the sensitive features, our FIF formulation cannot attribute any bias to the any other subset of  features, which leads to $ w_{\mathbf{S}} = 0 $ for all $\mathbf{S}$.

\textit{Expressing FIF in terms of the variance decomposition over a subset of features allows us to import and extend well-studied techniques of GSA to perform FIF estimation}, which we elaborate on Section~\ref{fairness_fairXplainer_sec:fairxplainer}.