### 2023-02-13 chapters/acknowledgements.tex
ack todo

\begin{acknowledgments}

\todo{Incomplete}With utmost humility, I would like to take this opportunity to express my sincere gratitude to everyone who has supported me throughout my PhD journey.

I come from a humble background in Bangladesh and have always had a deep passion for math, science, and computer science. My PhD journey has been a process of self-reflection and growth, from being an inattentive undergraduate student to becoming a more dedicated PhD graduate. I have learned to accept challenges and setbacks, which has taught me valuable lessons and contributed to my eventual success.

I am humbly grateful to my advisor, Kuldeep S. Meel, who has played a vital role in my PhD. He has taught me how to be a more independent researcher and his leadership, passion, and dedication have been a true source of inspiration. I feel privileged to have been a part of his research lab and to have benefited from its thriving research culture.

I have also been blessed to work with exceptional collaborators, including Debabrota Basu, Dmitry Malioutov, Daniel Neider, Naheed Anjum Arafat, Jonathan Scarlet, and Lorenzo. They have generously shared their knowledge and experience with me and have helped me broaden my perspectives. I have come to understand that finding a meaningful research problem is a crucial step and I am grateful to have had the opportunity to collaborate and brainstorm with such insightful colleagues and friends.

I would also like to extend my heartfelt thanks to my family, especially my mother, sister, brother-in-law, and uncle for their constant support and encouragement. While I regret that my father could not be here to see this day, I am comforted by the thought that his blessings will always be with me. Finally, I would like to thank my wife, who has been my supportive partner and companion throughout this journey. Without her unwavering support, it would have been a much more difficult road to completion.

Thank you all for your support and encouragement. I am truly grateful.

\vspace{2em}
\noindent Best regards,\\
Bishwamittra Ghosh

\end{acknowledgments}