### 2023-02-14 tables/fairness/fvgm/sanity_check.tex

\begin{figure}
	\begin{center}
		\subfloat[]{\includegraphics[scale=0.5]{figures/fairness/fvgm/synthetic_example_BN}\label{fairness_fvgm_fig:synthetic_bn}}

		\subfloat[]{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_ppv_max_PPV_LR}\label{fairness_fvgm_fig:synthetic_ppv_max_LR}}
		\subfloat[]{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_ppv_min_PPV_LR}\label{fairness_fvgm_fig:synthetic_ppv_min_LR}}

		\subfloat[]{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_sample_size_DI_LR}\label{fairness_fvgm_fig:synthetic_sample_size_LR_DI}}
		\subfloat[]{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_sample_size_SPD_LR}\label{fairness_fvgm_fig:synthetic_sample_size_LR_SPD}}
	\end{center}

	\caption[Accuracy of {\fvgm} on logistic regression classifier]{
		Measuring accuracy of different fairness verifiers for Example~\ref{fvgm_example:intro} on logistic regression classifier.}
	\label{fairness_fvgm_fig:synthetic_results}
\end{figure}

\begin{figure}
	\begin{center}
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_ppv_max_PPV_SVM}}
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_ppv_min_PPV_SVM}}\\
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_sample_size_DI_SVM}}
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_sample_size_SPD_SVM}}

	\end{center}

	\caption[Accuracy of {\fvgm} on SVM classifier]{
		Measuring accuracy of different fairness verifiers for Example~\ref{fvgm_example:intro} on SVM classifier.}
	\label{fairness_fvgm_fig:synthetic_results_SVM}
\end{figure}