### 2023-02-19 chapters/fairness/fvgm/conclusion.tex

\section{Chapter Summary}
In this chapter, we propose {\fvgm}, an efficient fairness verification framework for linear classifiers based on a novel stochastic subset-sum problem. {\fvgm} encodes a graphical model of feature-correlations, represented as a Bayesian Network, and computes multiple group and causal fairness metrics accurately. We experimentally demonstrate that {\fvgm} is not only more accurate and scalable than the existing verifiers but also applicable in practical fairness tasks, such as verifying fairness attacks and enhancing algorithms, and computing the fairness influence functions. In the next chapter, we elaborate on fairness influence functions as a framework to explaining group fairness metrics in machine learning.
\begin{comment}
	%As a future work, we aim to design fairness-enhancing algorithms certified by fairness verifiers, such as {\fvgm}. Since {\fvgm} serves as an accurate and scalable fairness verifier for linear classifiers, it will be interesting to design such verifiers for other machine learning models.
\end{comment}