### 2023-03-03 chapters/fairness/fvgm/conclusion.tex
minor edits

\section{Chapter Summary}
We discuss {\fvgm}, an efficient fairness verification framework for linear classifiers based on a novel stochastic subset-sum problem. {\fvgm} encodes a graphical model of feature-correlations, represented as a Bayesian Network, and computes multiple group and causal fairness metrics accurately. We experimentally demonstrate that {\fvgm} is  more accurate and scalable than the existing verifiers.
% but also applicable in practical fairness tasks, such as verifying fairness attacks and enhancing algorithms, and computing the fairness influence functions. 
% In the next chapter, we elaborate on fairness influence functions as a framework to explaining group fairness metrics in machine learning.
\begin{comment}
	%As a future work, we aim to design fairness-enhancing algorithms certified by fairness verifiers, such as {\fvgm}. Since {\fvgm} serves as an accurate and scalable fairness verifier for linear classifiers, it will be interesting to design such verifiers for other machine learning models.
\end{comment}