### 2023-07-18 chapters/acknowledgements.tex
acknowledgement added

\begin{acknowledgments}

I would like to express my sincere gratitude to everyone who has supported me throughout my PhD. This has been a journey of self-discovery, patience, and continuous effort to surpass myself day by day. I am fortunate to have met amazing researchers around the world, communicating ideas and learning from them. Reflecting on Socrates' famous words, ``Know Thyself'', my PhD has been a journey of understanding my limitations and fostering a mindset of overcoming shortcomings while enjoying the pursuit of knowledge. At this moment, I truly understand that the outcome of my PhD goes beyond the thesis itself; it encompasses my personal growth and development.

I come from a humble background in Bangladesh, where my passion for math, science, and computer science has always been a driving force. Today, I am immensely grateful to announce that my dream of pursuing a PhD in computer science has become a reality. I am deeply thankful to the brilliant minds I have had the privilege to encounter throughout my doctoral journey.

I am immensely grateful to my advisor, Dr. Kuldeep S. Meel, for playing a crucial role in my PhD journey. Under his guidance, I have learned the importance of being an independent researcher and have been provided with numerous opportunities to achieve this goal. I owe my confidence as a researcher to Kuldeep's unwavering support in connecting me with exceptional researchers through conferences, workshops, internships, and research visits. His leadership, passion, and dedication have truly inspired me. I consider myself privileged to have been a part of his research lab and to have experienced the vibrant research culture it offers. I vividly recall the early semesters in NUS, where Kuldeep would spend hours discussing with the research group, creating an environment that fostered growth and learning. As a new PhD student, his supervision boosted my confidence and enabled me to flourish in the doctoral program. Kuldeep's sharp intellect and innovative thinking have not only made me enjoy graduate school but have also facilitated my personal growth. I am grateful for his relentless guidance, which has not only enhanced my skills as a researcher but also shaped me into a better person.

I feel incredibly fortunate to have had the opportunity to collaborate with exceptional individuals throughout my research journey. Debabrota Basu, Dmitry Malioutov, Krisnha P. Gummadi, Daniel Neider, Arijit Khan, Jonathan Scarlet, Chandrika Bhardwaj, Vijay Saraswat, Naheed Anjum Arafat, and Lorenzo Ciampiconi are among the remarkable collaborators who have generously shared their knowledge and experience with me. I come to understand that finding a meaningful research problem is a crucial step and I am grateful to have had the opportunity to collaborate and brainstorm with such insightful collaborators.

I would like to express my sincere appreciation to my friends in Singapore who have provided invaluable support throughout my journey. Neamul Kabir, Sanjay Saha, Amlan Saha, Rafiul Islam, Zubair Faruqui, Mohimenul Mahi, Ramzan Miah, Muhim Muktadir Zim, and Shayok Ghosh have been pillars of support and encouragement. I am also grateful for the exceptional colleagues who have made a significant impact on my professional growth. Priyanka Golia, Teodora Baluta, Yash Pote, Arijit Shaw, Suwei Yang, Jiong Yang, Mate Soos, Tim van Bremen, Anna Latour, Yacine Izza, Gunjan Kumar, Paulius Dilkas, Shubham Sharma, Rahul Gupta, and Zhanzhong Pang have been sources of inspiration and collaboration. I truly believe that it takes an entire village to achieve something incredible, and I am fortunate to have such a supportive community. Additionally, I extend my thanks to all the staff members in SOC, NUS for their exceptional efficiency in addressing students' needs.

I would like to express my heartfelt gratitude to my family for their support and encouragement. My parents, sister, brother-in-law, and uncle have been constant pillars of strength throughout my journey. While I deeply regret that my father is not here with me today, I find solace in the belief that his blessings will always guide me. I am also immensely grateful to my wife, Dola Ghosh, who has been my unwavering source of support and companionship throughout this challenging journey. Her presence has made the road to completion much smoother, and I am grateful for her love and encouragement. Additionally, I would like to extend a special thank you to my childhood friends, Kalyan Roy, Sukanto Kundu, Nayem Ul Haque, and Niranjan Chandra Roy. Their friendship has been a constant source of strength and support throughout my life.

\begin{flushright}
	\vspace{1em}
	Bishwamittra Ghosh
\end{flushright}

\end{acknowledgments}